---
title: "The Office - Solution"
author: "Mine Çetinkaya-Rundel"
format: html
editor: visual
---

```{r}
#| label: load-packages
#| warning: false
#| message: false

library(tidyverse)
library(tidymodels)
library(schrute)
library(lubridate)
```

Use `theoffice` data from the [**schrute**](https://bradlindblad.github.io/schrute/) package to predict IMDB scores for episodes of The Office.

```{r}
glimpse(theoffice)
```

Fix `air_date` for later use.

```{r}
theoffice <- theoffice %>%
  mutate(air_date = ymd(as.character(air_date)))
```

We will

-   engineer features based on episode scripts
-   train a model
-   perform cross validation
-   make predictions

Note: The episodes listed in `theoffice` don't match the ones listed in the data we used in the [cross validation lesson](https://ids-s1-20.github.io/slides/week-10/w10-d02-cross-validation/w10-d02-cross-validation.html).

```{r}
theoffice %>%
  distinct(season, episode)
```

### Exercise 1 - Calculate the percentage of lines spoken by Jim, Pam, Michael, and Dwight for each episode of The Office.

```{r}
#| label: lines

office_lines <- theoffice %>%
  group_by(season, episode) %>%
  mutate(
    n_lines = n(),
    lines_jim = sum(character == "Jim") / n_lines,
    lines_pam = sum(character == "Pam") / n_lines,
    lines_michael = sum(character == "Michael") / n_lines,
    lines_dwight = sum(character == "Dwight") / n_lines,
  ) %>%
  ungroup() %>%
  select(season, episode, episode_name, contains("lines_")) %>%
  distinct(season, episode, episode_name, .keep_all = TRUE)
```

### Exercise 2 - Identify episodes that touch on Halloween, Valentine's Day, and Christmas.

```{r}
#| label: special-episodes

theoffice <- theoffice %>%
  mutate(text = tolower(text))

halloween_episodes <- theoffice %>%
  filter(str_detect(text, "halloween")) %>% 
  count(episode_name) %>%
  filter(n > 1) %>%
  mutate(halloween = 1) %>%
  select(-n)

valentine_episodes <- theoffice %>%
  filter(str_detect(text, "valentine")) %>% 
  count(episode_name) %>%
  filter(n > 1) %>%
  mutate(valentine = 1) %>%
  select(-n)

christmas_episodes <- theoffice %>%
  filter(str_detect(text, "christmas")) %>% 
  count(episode_name) %>%
  filter(n > 1) %>%
  mutate(christmas = 1) %>%
  select(-n)
```

### Exercise 3 - Put together a modeling dataset that includes features you've engineered. Also add an indicator variable called `michael` which takes the value `1` if Michael Scott (Steve Carrell) was there, and `0` if not. Note: Michael Scott (Steve Carrell) left the show at the end of Season 7.

```{r}
#| label: office-df

office_df <- theoffice %>%
  select(season, episode, episode_name, imdb_rating, total_votes, air_date) %>%
  distinct(season, episode, .keep_all = TRUE) %>%
  left_join(halloween_episodes, by = "episode_name") %>% 
  left_join(valentine_episodes, by = "episode_name") %>% 
  left_join(christmas_episodes, by = "episode_name") %>% 
  replace_na(list(halloween = 0, valentine = 0, christmas = 0)) %>%
  mutate(michael = if_else(season > 7, 0, 1)) %>%
  mutate(across(halloween:michael, as.factor)) %>%
  left_join(office_lines, by = c("season", "episode", "episode_name"))
```

### Exercise 4 - Split the data into training (75%) and testing (25%).

```{r}
#| label: split

set.seed(1122)
office_split <- initial_split(office_df)
office_train <- training(office_split)
office_test <- testing(office_split)
```

### Exercise 5 - Specify a linear regression model.

```{r}
#| label: model

office_mod <- linear_reg() %>%
  set_engine("lm")
```

### Exercise 6 - Create a recipe that updates the role of `episode_name` to not be a predictor, removes `air_date` as a predictor, and removes all zero variance predictors.

```{r}
#| label: recipe

office_rec <- recipe(imdb_rating ~ ., data = office_train) %>%
  update_role(episode_name, new_role = "id") %>%
  step_rm(air_date) %>%
  step_dummy(all_nominal(), -episode_name) %>%
  step_zv(all_predictors())
```

### Exercise 7 - Build a workflow for fitting the model specified earlier and using the recipe you developed to preprocess the data.

```{r}
#| label: workflow

office_wflow <- workflow() %>%
  add_model(office_mod) %>%
  add_recipe(office_rec)
```

### Exercise 8 - Fit the model to training data and interpret a couple of the slope coefficients.

```{r}
#| label: fit

office_fit <- office_wflow %>%
  fit(data = office_train)

tidy(office_fit)
```

### Exercise 9 - Perform 5-fold cross validation and view model performance metrics.

```{r}
#| label: cv
#| message: false
#| error: true

set.seed(345)
folds <- vfold_cv(office_train, v = 5)
folds

set.seed(456)
office_fit_rs <- office_wflow %>%
  fit_resamples(folds)

collect_metrics(office_fit_rs)
```

### Exercise 10 - Use your model to make predictions for the testing data and calculate the RMSE. Also use the model developed in the [cross validation lesson](https://ids-s1-20.github.io/slides/week-10/w10-d02-cross-validation/w10-d02-cross-validation.html) to make predictions for the testing data and calculate the RMSE as well. Which model did a better job in predicting IMDB scores for the testing data?

#### New model

```{r}
#| label: new-model

office_test_pred <- predict(office_fit, new_data = office_test) %>%
  bind_cols(office_test %>% select(imdb_rating, episode_name))

rmse(office_test_pred, truth = imdb_rating, estimate = .pred)
```

#### Old model

```{r}
#| label: old-model

office_mod_old <- linear_reg() %>%
  set_engine("lm")

office_rec_old <- recipe(imdb_rating ~ season + episode + total_votes + air_date, data = office_train) %>%
  # extract month of air_date
  step_date(air_date, features = "month") %>%
  step_rm(air_date) %>%
  # make dummy variables of month 
  step_dummy(contains("month")) %>%
  # remove zero variance predictors
  step_zv(all_predictors())

office_wflow_old <- workflow() %>%
  add_model(office_mod_old) %>%
  add_recipe(office_rec_old)

office_fit_old <- office_wflow_old %>%
  fit(data = office_train)

tidy(office_fit_old)

office_test_pred_old <- predict(office_fit_old, new_data = office_test) %>%
  bind_cols(office_test %>% select(imdb_rating, episode_name))

rmse(office_test_pred_old, truth = imdb_rating, estimate = .pred)
```
