---
title: "UN Votes"
author: "Mine Çetinkaya-Rundel"
date: "`r Sys.Date()`"
output: 
  html_document:
    toc: true
    toc_float: true
---

Let's take a look at the voting history of countries in the United Nations 
General Assembly. We will be using data from the __unvotes__ package. 
Additionally, we will make use of the __tidyverse__ and __lubridate__ packages 
for the analysis, and the __DT__ package for interactive display of tabular 
output.

```{r load-packages, warning=FALSE, message=FALSE}
library(unvotes)
library(tidyverse)
library(lubridate)
library(DT)
```

## Data

The unvotes package provides three datasets we can work with: `un_roll_calls`, 
`un_roll_call_issues`, and `un_votes`. Each of these datasets contains a 
variable called `rcid`, the roll call id, which can be used as a unique 
identifier to join them with each other.

- The `un_votes` dataset provides information on the voting history of the 
United Nations General Assembly. It contains one row for each country-vote pair.

```{r view-un_votes}
un_votes
```

- The `un_roll_calls` dataset contains information on each roll call vote of 
the United Nations General Assembly.

```{r view-un_roll_calls}
un_roll_calls
```

- The `un_roll_call_issues` dataset contains (topic) classifications of roll 
call votes of the United Nations General Assembly. Many votes had no topic, 
and some have more than one.

```{r view-un_roll_call_issues}
un_roll_call_issues
```

## Analysis

First, let's take a look at how often each country voted "Yes" on a resolution 
in each year. We'll visualize the results, so let's pick a few countries of 
interest first,

```{r country-list}
country_list <- c("United States of America", "Russian Federation", 
                  "Mexico", "Turkey")
```

and focus our analysis on them.

```{r plot-yearly-yes, fig.width=10, warning=FALSE}
un_votes %>%
  filter(country %in% country_list) %>%
  inner_join(un_roll_calls, by = "rcid") %>%
  group_by(year = year(date), country) %>%
  summarize(
    votes = n(),
    percent_yes = mean(vote == "yes")
    ) %>%
  ggplot(mapping = aes(x = year, y = percent_yes, color = country)) +
    geom_line() +
    labs(
      title = "Percentage of 'Yes' votes in the UN General Assembly",
      subtitle = "1946 to 2015",
      y = "% Yes",
      x = "Year",
      color = "Country"
    )
```

Next, let's create a visualization that we create a visualization that displays 
how the voting record of the United States changed over time on a variety of 
issues, and compares it to another country. The other country we'll display is 
Turkey.

```{r plot-yearly-yes-issue, fig.width=10, warning=FALSE}
un_votes %>%
  filter(country %in% c("United States of America", "Turkey")) %>%
  inner_join(un_roll_calls, by = "rcid") %>%
  inner_join(un_roll_call_issues, by = "rcid") %>%
  group_by(country, year = year(date), issue) %>%
  summarize(
    votes = n(),
    percent_yes = mean(vote == "yes")
    ) %>%
  filter(votes > 5) %>%  # only use records where there are more than 5 votes
  ggplot(mapping = aes(x = year, y = percent_yes, color = country)) +
    geom_point() +
    geom_smooth(method = "loess", se = FALSE) +
    facet_wrap(~ issue) +
    labs(
      title = "Percentage of 'Yes' votes in the UN General Assembly",
      subtitle = "1946 to 2015",
      y = "% Yes",
      x = "Year",
      color = "Country"
    )
```

We can easily change which countries are being plotted by changing which 
countries the code above `filter`s for. Note that the country name should be 
spelled and capitalized exactly the same way as it appears in the data. See 
the [Appendix](#appendix) for a list of the countries in the data.

## References

1. David Robinson (2017). unvotes: United Nations General Assembly Voting Data. 
R package version 0.2.0. https://CRAN.R-project.org/package=unvotes.
2. Erik Voeten "Data and Analyses of Voting in the UN General Assembly" 
Routledge Handbook of International Organization, edited by Bob Reinalda 
(published May 27, 2013).
3. Much of the analysis has been modeled on the examples presented in the 
[unvotes package vignette](https://cran.r-project.org/web/packages/unvotes/vignettes/unvotes.html).

## Appendix {#appendix}

Below is a list of countries in the dataset:

```{r list-countries}
un_votes %>% 
  select(country) %>%
  arrange(country) %>% 
  distinct() %>%
  datatable()
```
