---
title: "Linear models with multiple predictors <br> `r emo::ji('woman_juggling')`"
author: ""
output:
  xaringan::moon_reader:
    css: "../slides.css"
    lib_dir: libs
    nature:
      ratio: "16:9"
      highlightLines: true
      highlightStyle: solarized-light
      countIncrementalSlides: false
---

```{r child = "../setup.Rmd"}
```

```{r packages, echo=FALSE, message=FALSE, warning=FALSE}
library(tidyverse)
library(scales)
library(broom)
library(here)
library(plotly)
library(widgetframe)
```

class: middle

# The linear model with multiple predictors

---

## The data

```{r load-pp, message=FALSE}
pp <- read_csv(
  "data/paris-paintings.csv", 
  na = c("n/a", "", "NA")
  )
```

---

## Data transformation

```{r transform-pp, message=FALSE}
pp <-pp %>%
  mutate(
    log_price = log(price),
    artistliving = factor(artistliving)
  )
```

---

## Multiple predictors

- Response variable: `log_price` = log_price
- Explanatory variables: Width and height

```{r model-price-width-height}
m_wi_hgt <- lm(log_price ~ Width_in + Height_in, data = pp)
tidy(m_wi_hgt)
```

---

##  Linear model with multiple predictors

```{r model-price-width-height-tidy, echo=FALSE}
tidy(m_wi_hgt)
```

<br>

$$\widehat{log_price} = 4.77 + 0.0269~width - 0.0133~height$$

---

## Visualizing models with multiple predictors

```{r plotly, echo=FALSE, warning=FALSE, cache=TRUE}
p <- plot_ly(pp, x = ~Width_in, y = ~Height_in, z = ~log_price,
        marker = list(size = 3,
                       color = "lightgray",
                       alpha = 0.5,
                       line = list(color = "gray",
                                   width = 2))) %>%
  add_markers() %>%
  plotly::layout(scene = list(xaxis = list(title = "Width (in)"),
                      yaxis = list(title = "Height (in)"),
                      zaxis = list(title = "log_price"))) %>%
  config(displayModeBar = FALSE)
frameWidget(p, width = "100%", height = "100%")
```

---

class: middle

# Exploration: Price, surface area, and living artist

---

## Price, surface area, and living artist

- Explore the relationship between price of paintings and surface area, conditioned 
on whether or not the artist is still living
- First visualize and explore, then model
- But first, prep the data

.midi[
```{r}
pp <- pp %>%
  mutate(artistliving = if_else(artistliving == 0, "Deceased", "Living"))

pp %>%
  count(artistliving)
```
]

---

## Typical surface area

.question[
What is the typical surface area for paintings?
]

```{r viz-surf-artistliving, echo=FALSE,warning=FALSE, out.width="80%"}
ggplot(data = pp, 
       mapping = aes(x = Surface, fill = artistliving)) +
  geom_histogram(binwidth = 500) + 
  facet_grid(artistliving ~ .) +
  theme_minimal() +
  theme(legend.position = "none") +
  labs(y = "") +
  geom_vline(xintercept = 1000) +
  geom_vline(xintercept = 5000, linetype = "dashed", color = "gray")
```

--

Less than 1000 square inches (~ 80cm x 80cm). There are very few paintings that have surface area above 5000 square inches.

---

## Narrowing the scope

For simplicity let's focus on the paintings with `Surface < 5000`:

```{r surf-lt-5000, out.width="90%"}
pp_Surf_lt_5000 <- pp %>%
  filter(Surface < 5000)
```

```{r viz-surf-lt-5000-artistliving, echo=FALSE, warning=FALSE, out.width="80%"}
ggplot(data = pp_Surf_lt_5000, 
       mapping = aes(y = log_price, x = Surface, 
                     color = artistliving)) +
  geom_point(alpha = 0.3) +
  labs(color = "Living artist") +
  theme_minimal()
```

---

## Facet to get a better look

```{r viz-surf-lt-5000-artistliving-facet, echo=FALSE, warning=FALSE, out.width="80%"}
ggplot(data = pp_Surf_lt_5000, 
       mapping = aes(y = log_price, x = Surface, 
                     color = artistliving)) +
  geom_point(alpha = 0.3) +
  facet_wrap(~artistliving) +
  labs(color = "Living artist") +
  theme_minimal()
```

---

## Two ways to model

- **Main effects:** Assuming relationship between surface and logged price 
**does not vary** by whether or not the artist is living.
- **Interaction effects:** Assuming relationship between surface and logged 
price **varies** by whether or not the artist is living.

---

## Interacting explanatory variables

- Including an interaction effect in the model allows for different slopes, i.e. 
nonparallel lines.
- This implies that the regression coefficient for an explanatory variable would 
change as another explanatory variable changes.
- This can be accomplished by adding an interaction variable: the product of two 
explanatory variables.

---

class: middle

# Side-step: Weights of books

---

.question[
Suppose we want to predict weight of books from their volume and cover type 
(hardback vs. paperback). Do you think a model with main effects or 
interaction effects is more appropriate? Explain your reasoning.

**Hint:** Main effects would mean rate at which weight changes as volume 
increases would be the same for hardback and paperback books and interaction 
effects would mean the rate at which weight changes as volume 
increases would be different for hardback and paperback books.
]

```{r echo=FALSE}
countdown(minutes = 3)
```

---

The `allbacks` data frame gives measurements on the volume and weight of 15 books, some of which are softback and some of which are hardback.

.small[
```{r}
library(DAAG)
as_tibble(allbacks)
```
]

.footnote[
The bookshelf of J. H. Maindonald at Australian National University.
]

---

```{r out.width="80%"}
ggplot(allbacks, aes(x = volume, y = weight, color = cover)) +
  geom_point(alpha = 0.7) +
  theme_minimal()
```

---

```{r echo=FALSE, out.width="55%"}
m_main <- lm(weight ~ volume + cover, data = allbacks)
m_main_aug <- augment(m_main)

ggplot() +
  geom_point(allbacks, 
             mapping = aes(x = volume, y = weight, color = cover), alpha = 0.7) +
  geom_line(m_main_aug, 
              mapping = aes(x = volume, y = .fitted, color = cover)) +
  theme_minimal() +
  labs(title = "Main effects, parallel slopes", 
       subtitle = "weight-hat = volume + cover")
```


```{r echo=FALSE, out.width="55%"}
m_int <- lm(weight ~ volume + cover + volume*cover, data = allbacks)
m_int_aug <- augment(m_int)

ggplot() +
  geom_point(data = allbacks, 
             mapping = aes(x = volume, y = weight, color = cover), alpha = 0.7) +
  geom_line(data = m_int_aug, 
            mapping = aes(x = volume, y = .fitted, color = cover)) +
  theme_minimal() +
  labs(title = "Interaction effects, not parallel slopes", 
       subtitle = "weight-hat = volume + cover + volume * cover")
```

---

## In pursuit of Occam's razor

- Occam's Razor states that among competing hypotheses that predict equally 
well, the one with the fewest assumptions should be selected.
- Model selection follows this principle.
- We only want to add another variable to the model if the addition of that
variable brings something valuable in terms of predictive power to the model.
- In other words, we prefer the simplest best model, i.e. **parsimonious** model.

---

.question[
Visually, which of the two models is preferable under Occam's razor?
]

```{r echo=FALSE, out.width="55%"}
m_main <- lm(weight ~ volume + cover, data = allbacks)
m_main_aug <- augment(m_main)

ggplot() +
  geom_point(allbacks, 
             mapping = aes(x = volume, y = weight, color = cover), alpha = 0.7) +
  geom_line(m_main_aug, 
              mapping = aes(x = volume, y = .fitted, color = cover)) +
  theme_minimal() +
  labs(title = "Main effects, parallel slopes", 
       subtitle = "weigmainht-hat = volume + cover")
```


```{r echo=FALSE, out.width="55%"}
m_int <- lm(weight ~ volume + cover + volume*cover, data = allbacks)
m_int_aug <- augment(m_int)

ggplot() +
  geom_point(data = allbacks, 
             mapping = aes(x = volume, y = weight, color = cover), alpha = 0.7) +
  geom_line(data = m_int_aug, 
            mapping = aes(x = volume, y = .fitted, color = cover)) +
  theme_minimal() +
  labs(title = "Interaction effects, not parallel slopes", 
       subtitle = "weight-hat = volume + cover + volume * cover")
```

---

## R-squared

- $R^2$ is the percentage of variability in the response variable explained by 
the regression model.

```{r}
glance(m_main)$r.squared
glance(m_int)$r.squared
```

--
- Clearly the model with interactions has a higher $R^2$.

--
- However using $R^2$ for model selection in models with multiple explanatory 
variables is not a good idea as $R^2$ increases when **any** variable is added 
to the model.

---

## Adjusted R-squared

... a (more) objective measure for model selection

- Adjusted $R^2$ doesn't increase if the new variable does not provide any new 
informaton or is completely unrelated, as it applies a penalty for number of 
variables included in the model.
- This makes adjusted $R^2$ a preferable metric for model selection in multiple
regression models.

---

## Comparing models

.pull-left[
```{r}
glance(m_main)$r.squared
glance(m_int)$r.squared
```
]
.pull-right[
```{r}
glance(m_main)$adj.r.squared
glance(m_int)$adj.r.squared
```
]

--

.small[
```{r}
# Is R-sq higher for int model?
glance(m_int)$r.squared > glance(m_main)$r.squared 
# Is R-sq adj. higher for int model?
glance(m_int)$adj.r.squared > glance(m_main)$adj.r.squared
```
]

---

class: middle

# Back to exploration: Price, surface area, and living artist

---

## Two ways to model

- **Main effects:** Assuming relationship between surface and logged price 
**does not vary** by whether or not the artist is living.
- **Interaction effects:** Assuming relationship between surface and logged 
price **varies** by whether or not the artist is living.

.pull-left[
```{r viz-main-effects, out.width="80%", echo=FALSE}
m_main <- lm(log_price ~ Surface + artistliving, data = pp_Surf_lt_5000)
m_main_aug <- augment(m_main)
ggplot(data = m_main_aug, 
       mapping = aes(y = log_price, x = Surface, color = artistliving)) +
  geom_point(alpha = 0.3) +
  geom_line(aes(y = .fitted), size = 1.2) +
  labs(x = "Surface", y = "log_price", color = "Living artist", 
       title = "Main effects") +
  theme_minimal()
```
]
.pull-right[
```{r viz-interaction-effects, out.width="80%", echo=FALSE}
ggplot(data = pp_Surf_lt_5000,
       mapping = aes(y = log_price, x = Surface, 
                     color = artistliving)) +
  geom_point(alpha = 0.3) +
  geom_smooth(method = "lm", se = FALSE) +
  labs(y = "log_price", color = "Living artist", title = "Interaction effects") +
  theme_minimal()
```
]

---

## Fit model with main effects

- Response variable: log_price
- Explanatory variables: Surface area and artist living (0/1 variable)

.midi[
```{r model-main-effects}
m_main <- lm(log_price ~ Surface + artistliving, 
             data = pp_Surf_lt_5000)
tidy(m_main)
```
]


--

$$ \widehat{log_price} = 4.88 + 0.000265~surface + 0.137~artistliving $$

---

## Solving the model

- Non-living artist: Plug in 0 for `artistliving`

$\widehat{log_price} = 4.88 + 0.000265~surface + 0.137 \times 0$  
$= 4.88 + 0.000265~surface$

--
- Living artist: Plug in 1 for `artistliving`

$\widehat{log_price} = 4.88 + 0.000265~surface + 0.137 \times 1$   
$= 5.017 + 0.000265~surface$

---

## Visualizing main effects

```{r out.width="60%", echo = FALSE}
ggplot(data = m_main_aug, 
       mapping = aes(y = log_price, x = Surface, color = artistliving)) +
  geom_point(alpha = 0.3) +
  geom_line(aes(y = .fitted)) +
  labs(x = "Surface", y = "log_price", color = "Living artist") +
  theme_minimal()
```

- **Same slope:** Rate of change in price as the surface area increases does 
not vary between paintings by living and non-living artists.
- **Different intercept:** Paintings by living artists are consistently more 
expensive than paintings by non-living artists.

---

## Interpreting main effects

.midi[
```{r exp-coefs}
tidy(m_main) %>% 
  mutate(exp_estimate = exp(estimate)) %>%
  select(term, estimate, exp_estimate)
```
]

- All else held constant, for each additional square inch in painting's surface area, the price of the painting is predicted, on average, to be higher by a factor of 1.
- All else held constant, paintings by a living artist are predicted, on average, to be higher by a factor of 1.15 compared to paintings by an artist who is no longer alive.
- Paintings that are by an artist who is not alive and that have a surface area of 0 square inches are predicted, on average, to be 132 livres.

---

## 

.question[
Why is our linear regression model different from what we got from `geom_smooth(method = "lm")`?
]

.pull-left[
```{r echo=FALSE, out.width="80%"}
ggplot(pp_Surf_lt_5000, aes(x = Surface, y = log_price, color = artistliving)) + 
  geom_point(alpha = 0.3) +
  geom_smooth(method = "lm") +
  labs(x = "Surface", y = "log_price", color = "Living artist") +
  theme_minimal()
```
]
.pull-right[
```{r viz-main-effects3, echo=FALSE, out.width="80%"}
m_pr <- lm(log_price ~ Surface + artistliving, data = pp_Surf_lt_5000)
m_pr_aug <- augment(m_pr)
ggplot(data = m_pr_aug, mapping = aes(y = log_price, x = Surface, color = artistliving)) +
  geom_point(alpha = 0.3) +
  geom_line(aes(y = .fitted)) +
  labs(x = "Surface", y = "log_price", color = "Living artist") +
  theme_minimal()
```
]

---

## What went ~wrong~ diferently?

- The way we specified our model only lets `artistliving` affect the intercept.
- Model implicitly assumes that paintings with living and deceased artists have the *same slope* and only allows for *different intercepts*.  

.question[
What seems more appropriate in this case?
+ Same slope and same intercept for both colors
+ Same slope and different intercept for both colors
+ Different slope and different intercept for both colors
]

---

## Interaction: surface * artist living

```{r out.width="80%", echo = FALSE}
m_int <- lm(log_price ~ Surface * artistliving, data = pp_Surf_lt_5000)
m_int_aug <- augment(m_int)
ggplot(data = m_int_aug, 
       mapping = aes(y = log_price, x = Surface, color = artistliving)) +
  geom_point(alpha = 0.3) +
  geom_line(aes(y = .fitted)) +
  labs(x = "Surface", y = "log_price", color = "Living artist") +
  theme_minimal()
```

---

## Fit model with interaction effects

- Response variable: log_price
- Explanatory variables: Surface area, artist living (0/1 variable), and 
their interaction

.midi[
```{r model-interaction-effects}
m_int <- lm(log_price ~ Surface + artistliving + 
              Surface * artistliving, 
            data = pp_Surf_lt_5000)
tidy(m_int)
```
]

---

## Linear model with interaction effects

.midi[
```{r model-interaction-effects-tidy, echo=FALSE}
tidy(m_int)
```
]

$$ \widehat{log_price} = 4.91 + 0.00021~surface - 0.126~artistliving $$
$$+ ~ 0.00048~surface \times artistliving $$

---

## Interpretation of interaction effects

- Rate of change in price as the surface area of the painting increases does 
vary between paintings by living and non-living artists (different slopes), 
- Some paintings by living artists are more expensive than paintings by
non-living artists, and some are not (different intercept).

.small[
.pull-left[
- Non-living artist: 
$\widehat{log_price} = 4.91 + 0.00021~surface$
$- 0.126 \times 0 + 0.00048~surface \times 0$
$= 4.91 + 0.00021~surface$
- Living artist: 
$\widehat{log_price} = 4.91 + 0.00021~surface$
$- 0.126 \times 1 + 0.00048~surface \times 1$
$= 4.91 + 0.00021~surface$
$- 0.126 + 0.00048~surface$
$= 4.784 + 0.00069~surface$
]
.pull-right[
```{r viz-interaction-effects2, out.width="80%", echo = FALSE}
ggplot(data = pp_Surf_lt_5000,
       aes(y = log_price, x = Surface, color = artistliving)) +
  geom_point(alpha = 0.3) +
  stat_smooth(method = "lm", fullrange = TRUE) +
  theme_minimal()
```
]
]

---

## Comparing models

It appears that adding the interaction actually increased adjusted $R^2$, so we 
should indeed use the model with the interactions.

```{r}
glance(m_main)$adj.r.squared
glance(m_int)$adj.r.squared
```

---

## Third order interactions

- Can you? Yes
- Should you? Probably not if you want to interpret these interactions in context
of the data.
