---
title: "Prediction and overfitting"
subtitle: "<br><br> Data Science in a Box"
author: "[datasciencebox.org](https://datasciencebox.org/)"
output:
  xaringan::moon_reader:
    css: ["../xaringan-themer.css", "../slides.css"]
    lib_dir: libs
    anchor_sections: FALSE
    nature:
      ratio: "16:9"
      highlightLines: true
      highlightStyle: solarized-light
      countIncrementalSlides: false
---

```{r child = "../setup.Rmd"}
```

```{r packages, echo = FALSE, message=FALSE, warning=FALSE}
library(tidyverse)
library(tidymodels)
library(openintro)
library(patchwork)
library(skimr)
set.seed(1234)
options(
  warnPartialMatchArgs = FALSE,
  warnPartialMatchAttr = FALSE, 
  warnPartialMatchDollar = FALSE,
  width = 100
)
```

class: middle

# Prediction

---

## Goal: Building a spam filter

- Data: Set of emails and we know if each email is spam/not and other features 
- Use logistic regression to predict the probability that an incoming email is spam
- Use model selection to pick the model with the best predictive performance

--
- Building a model to predict the probability that an email is spam is only half of the battle! We also need a decision rule about which emails get flagged as spam (e.g. what probability should we use as out cutoff?)

--
- A simple approach: choose a single threshold probability and any email that exceeds that probability is flagged as spam

---

## A multiple regression approach

.panelset[
.panel[.panel-name[Output]
.small[
```{r ref.label = "spam-mlr", echo = FALSE, warning = FALSE}
```
]
]
.panel[.panel-name[Code]
```{r spam-mlr, results = "hide"}
logistic_reg() %>%
  set_engine("glm") %>%
  fit(spam ~ ., data = email, family = "binomial") %>%
  tidy() %>%
  print(n = 22)
```
]
]

---

## Prediction

- The mechanics of prediction is **easy**:
  - Plug in values of predictors to the model equation
  - Calculate the predicted value of the response variable, $\hat{y}$

--
- Getting it right is **hard**!
  - There is no guarantee the model estimates you have are correct
  - Or that your model will perform as well with new data as it did with your sample data

---

## Underfitting and overfitting

```{r echo=FALSE, out.width="70%", warning = FALSE}
lm_fit <- linear_reg() %>%
  set_engine("lm") %>%
  fit(y4 ~ x2, data = association)

loess_fit <- loess(y4 ~ x2, data = association)

loess_overfit <- loess(y4 ~ x2, span = 0.05, data = association)

association %>%
  select(x2, y4) %>%
  mutate(
    Underfit = augment(lm_fit$fit) %>% select(.fitted) %>% pull(),
    OK       = augment(loess_fit) %>% select(.fitted) %>% pull(),
    Overfit  = augment(loess_overfit) %>% select(.fitted) %>% pull(),
  ) %>%
  pivot_longer(
    cols      = Underfit:Overfit,
    names_to  = "fit",
    values_to = "y_hat"
  ) %>%
  mutate(fit = fct_relevel(fit, "Underfit", "OK", "Overfit")) %>%
  ggplot(aes(x = x2)) +
  geom_point(aes(y = y4), color = "darkgray") +
  geom_line(aes(y = y_hat, group = fit, color = fit), size = 1) +
  labs(x = NULL, y = NULL, color = NULL) +
  scale_color_viridis_d(option = "plasma", end = 0.7)
```

---

## Spending our data

- Several steps to create a useful model: parameter estimation, model selection, performance assessment, etc.

- Doing all of this on the entire data we have available can lead to **overfitting**

- Allocate specific subsets of data for different tasks, as opposed to allocating the largest possible amount to the model parameter estimation only (which is what we've done so far)

---

class: middle

# Splitting data

---

## Splitting data

- **Training set:**
  - Sandbox for model building 
  - Spend most of your time using the training set to develop the model
  - Majority of the data (usually 80%)
  
- **Testing set:**
  - Held in reserve to determine efficacy of one or two chosen models
  - Critical to look at it once, otherwise it becomes part of the modeling process
  - Remainder of the data (usually 20%)
  
---

## Performing the split

```{r}
# Fix random numbers by setting the seed 
# Enables analysis to be reproducible when random numbers are used 
set.seed(1116)

# Put 80% of the data into the training set 
email_split <- initial_split(email, prop = 0.80)

# Create data frames for the two sets:
train_data <- training(email_split)
test_data  <- testing(email_split)
```

---

## Peek at the split

.small[
.pull-left[
```{r}
glimpse(train_data)
```
]
.pull-right[
```{r}
glimpse(test_data)
```
]
]

---

class: middle

# Modeling workflow

---

## Fit a model to the training dataset

```{r}
email_fit <- logistic_reg() %>%
  set_engine("glm") %>%
  fit(spam ~ ., data = train_data, family = "binomial")
```

---

## Categorical predictors

```{r echo=FALSE, out.width="75%", fig.width=10}
factor_predictors <- train_data %>%
  select(where(is.factor), -spam) %>%
  names()

p_to_multiple <- ggplot(train_data, aes(x = to_multiple, fill = spam)) +
  geom_bar() +
  scale_fill_manual(values = c("#E48957", "#CA235F"))

p_from <- ggplot(train_data, aes(x = from, fill = spam)) +
  geom_bar() +
  scale_fill_manual(values = c("#E48957", "#CA235F"))

p_sent_email <- ggplot(train_data, aes(x = sent_email, fill = spam)) +
  geom_bar() +
  scale_fill_manual(values = c("#E48957", "#CA235F"))

p_winner <- ggplot(train_data, aes(x = winner, fill = spam)) +
  geom_bar() +
  scale_fill_manual(values = c("#E48957", "#CA235F"))

p_format <- ggplot(train_data, aes(x = format, fill = spam)) +
  geom_bar() +
  scale_fill_manual(values = c("#E48957", "#CA235F"))

p_re_subj <- ggplot(train_data, aes(x = re_subj, fill = spam)) +
  geom_bar() +
  scale_fill_manual(values = c("#E48957", "#CA235F"))

p_urgent_subj <- ggplot(train_data, aes(x = urgent_subj, fill = spam)) +
  geom_bar() +
  scale_fill_manual(values = c("#E48957", "#CA235F"))

p_number <- ggplot(train_data, aes(x = number, fill = spam)) +
  geom_bar() +
  scale_fill_manual(values = c("#E48957", "#CA235F"))

p_to_multiple + p_from + p_sent_email + p_winner + p_format + p_re_subj + p_urgent_subj + p_number +
  plot_layout(ncol = 4, guides = "collect") & 
  theme(axis.title.y = element_blank())
```

---

## `from` and `sent_email`

.pull-left[
- `from`: Whether the message was listed as from anyone (this is usually set by default for regular outgoing email)

```{r}
train_data %>%
  count(spam, from)
```
]
.pull-right[
- `sent_email`: Indicator for whether the sender had been sent an email in the last 30 days

```{r}
train_data %>%
  count(spam, sent_email)
```
]

---

## Numerical predictors

.small[
```{r echo=FALSE, highlight.output=c(14, 15)}
options(width = 100)
x <- train_data %>%
  group_by(spam) %>%
  select(where(is.numeric)) %>%
  skim_without_charts()

print(x, include_summary = FALSE)
```
]

---

## Fit a model to the training dataset

```{r}
email_fit <- logistic_reg() %>%
  set_engine("glm") %>%
  fit(spam ~ . - from - sent_email - viagra, data = train_data, family = "binomial") #<<
```

.small[
```{r}
email_fit
```
]

---

## Predict outcome on the testing dataset

```{r}
predict(email_fit, test_data)
```


---

## Predict probabilities on the testing dataset

```{r}
email_pred <- predict(email_fit, test_data, type = "prob") %>%
  bind_cols(test_data %>% select(spam, time))

email_pred
```

---

## A closer look at predictions

.pull-left-wide[
```{r highlight.output=c(6, 10)}
email_pred %>%
  arrange(desc(.pred_1)) %>%
  print(n = 10)
```
]

---

## Evaluate the performance

**Receiver operating characteristic (ROC) curve**<sup>+</sup> which plot true positive rate vs. false positive rate (1 - specificity)

.pull-left[
```{r roc, fig.show="hide"}
email_pred %>%
  roc_curve(
    truth = spam,
    .pred_1,
    event_level = "second"
  ) %>%
  autoplot()
```
]
.pull-right[
```{r ref.label="roc", echo=FALSE, out.width="100%"}
```
]

.footnote[
.small[
<sup>+</sup>Originally developed for operators of military radar receivers, hence the name.
]
]

---

## Evaluate the performance

Find the area under the curve:

.pull-left[
```{r}
email_pred %>%
  roc_auc(
    truth = spam,
    .pred_1,
    event_level = "second"
  )
```
]
.pull-right[
```{r ref.label="roc", echo=FALSE, out.width="100%"}
```
]
