---
title: "Fisheries"
author: "Mine Çetinkaya-Rundel"
date: "`r Sys.Date()`"
output: 
  html_document: 
    toc: yes
    toc_float: yes
---

```{r setup, include=FALSE}
library(tidyverse)
fisheries <- read_csv("data/fisheries.csv")
continents <- read_csv("data/continents.csv")
```

The code below brings you to the point in the slides where we left off...

```{r catch-up}
fisheries <- fisheries %>%
  filter(total > 100000) %>%
  left_join(continents) %>%
  mutate(
    continent = case_when(
      country == "Democratic Republic of the Congo" ~ "Africa",
      country == "Hong Kong"                        ~ "Asia",
      country == "Myanmar"                          ~ "Asia",
      TRUE                                          ~ continent
      ),
    aquaculture_perc = aquaculture / total
    )
```

**Note:** In each of these exercises you will need to set `eval=TRUE` when you're 
ready to run the code for that exercise.

### Exercise 1.

Calculate the mean aquaculture percentage (we'll call it `mean_ap` for short) 
for continents in the fisheries data using the `summarise()` function in dplyr. 
Note that the function for calculating the mean is `mean()` in R.

```{r fisheries-mean, eval=FALSE}
fisheries %>%                  # start with the fisheries data frame
  ___ %>%                      # group by continent
  ___(mean_ap = ___)           # calculate mean aquaculture
```

### Exercise 2.

Now expand your calculations to also calculate the minimum and maximum 
aquaculture percentage for continents in the fisheries data. Note that the 
functions for calculating minimum and maximum in R are `min()` and `max()`
respectively.

```{r fisheries-summary, eval=FALSE}
fisheries %>%                  # start with the fisheries data frame
  # and the rest of the code goes here         
```

### Exercise 3.

Create a new data frame called `fisheries_summary_continent` that calculates 
minimum, mean, and maximum aquaculture percentage for each continent in the 
fisheries data. 

```{r fisheries-summary-continent, eval=FALSE}
fisheries_summary_continent <- fisheries %>%
  # you can reuse code from Exercise 2 here                        
```

### Exercise 4.

Take the `fisheries_summary_continent` data frame and order the results in descending 
order of mean aquaculture percentage.

```{r fisheries-summary-continent-sorted, eval=FALSE}
fisheries_summary_continent %>%      # start with the fisheries_summary_continent data frame
  ___                                # order in descending order of mean_ap
```
