# Day one {#dayone}

You only get one first day of class, so start with something that excites 
students, teach the necessary evils later. This means getting to a meaningful, 
and hopefully interesting (for the students) data visualization as quickly 
as possible.

## What is data science?

Data science is an exciting discipline that allows you to turn raw data into 
understanding, insight, and knowledge. We're going to learn to do this in a 
`tidy` way -- more on that later!

## What is this course?

This is an introductory data science course designed for learners with no 
background in data science, programming, or statistics, but a willingness to 
learn in class and explore independently.

- Will we be doing computing? Yes.
- Is this an intro CS course? Not really, but many themes are shared.
- Is this an intro stat course? In a way, but it's not your high school 
statistics course.
- What computing language will we learn? R.
- Why not language X? We can discuss that over :coffee:.

## Data in the wild

I like starting off by showing a few examples of easy to follow but not so 
simple data analyses done in R, preferably presented along with the R code. 
This should be the type of analysis they could do for their final project. 
Blog posts are often good examples for these. In Spring 2018 I used the 
following as examples:

- [A year as told by fitbit](http://livefreeordichotomize.com/2017/12/27/a-year-as-told-by-fitbit/) by Nick Strayer
- [R-Ladies global tour](http://www.masalmon.eu/2017/10/06/globalrladiestour/) by Maelle Salmon
- [Text analysis of Trump's tweets confirms he writes only the (angrier) 
Android half](http://varianceexplained.org/r/trump-tweets/) by David Robinson

## Your turn: UN Votes

It's now time to let the students work on their first data visualization in R. 

Using the `unvotes` package, and a pre-populated R Markdown document on RStudio 
Cloud, they can create and modify the two multi-variate plots, visualizing the 
voting history of countries in the United Nations General Assembly.

```{r load-packages, warning=FALSE}
library(unvotes)
library(tidyverse)
library(lubridate)
```

We will narrow down the analysis to four countries: United States, Russian 
Federation, Mexico, and Turkey.

```{r country-list}
country_list <- c("United States of America", "Russian Federation", 
                  "Mexico", "Turkey")
```

First we take a look at how often each country voted "yes" on a resolution 
in each year.

```{r plot-yearly-yes, warning=FALSE}
un_votes %>%
  filter(country %in% country_list) %>%
  inner_join(un_roll_calls, by = "rcid") %>%
  group_by(year = year(date), country) %>%
  summarize(
    votes = n(),
    percent_yes = mean(vote == "yes")
    ) %>%
  ggplot(mapping = aes(x = year, y = percent_yes, color = country)) +
    geom_line() +
    labs(
      title = "Percentage of 'Yes' votes in the UN General Assembly",
      subtitle = "1946 to 2015",
      y = "% Yes",
      x = "Year",
      color = "Country"
    )
```

And then, we create a visualization that compares how the voting record of the 
United States changed over time on a variety of issues, and compare it to 
another country.

```{r plot-yearly-yes-issue, fig.height=4, fig.width=7, warning=FALSE}
un_votes %>%
  filter(country %in% c("United States of America", "Turkey")) %>%
  inner_join(un_roll_calls, by = "rcid") %>%
  inner_join(un_roll_call_issues, by = "rcid") %>%
  group_by(country, year = year(date), issue) %>%
  summarize(
    votes = n(),
    percent_yes = mean(vote == "yes")
    ) %>%
  filter(votes > 5) %>%  # only use records where there are more than 5 votes
  ggplot(mapping = aes(x = year, y = percent_yes, color = country)) +
    geom_point() +
    geom_smooth(method = "loess", se = FALSE) +
    facet_wrap(~ issue) +
    labs(
      title = "Percentage of 'Yes' votes in the UN General Assembly",
      subtitle = "1946 to 2015",
      y = "% Yes",
      x = "Year",
      color = "Country"
    )
```

At this point, instead of a formal introduction on R syntax, I recommend 
letting students change parameters passed to these functions, such as which 
countries are being plotted, and reknit the document to view the changes.

[TO DO: Find a good way to insert slides.]

Link to relevant slides: https://github.com/rstudio-education/datascience-box/tree/master/slides/p0_d01-welcome


