---
title: "Lab 11 - So what if you smoke when pregnant?"
subtitle: "Simulation based inference"
output: 
  tufte::tufte_html:
    tufte_variant: "envisioned"
    highlight: pygments
    css: ../lab.css
link-citations: yes
---

```{r include=FALSE}
library(tufte)
library(knitr)
options(
  htmltools.dir.version = FALSE, # for blogdown
  show.signif.stars = FALSE,     # for regression output
  digits = 2
  )
knitr::opts_chunk$set(eval = TRUE)
```

In 2004, the state of North Carolina released a large data set containing 
information on births recorded in this state. This data set is useful to 
researchers studying the relation between habits and practices of expectant 
mothers and the birth of their children. We will work with a random sample of 
observations from this data set.

## Getting started

### Packages

In this lab we will work with the **tidyverse**, **infer**, and **openintro** packages. We can install and load them with the following:

```{r load-packages, message=FALSE}
library(tidyverse) 
library(infer)
library(openintro)
```

## Housekeeping

### Git configuration / password caching

Configure your Git user name and email. If you cannot remember the instructions, refer to an earlier lab. Also remember that you can cache your password for a limited amount of time.

### Project name

Update the name of your project to match the lab's title.

## Warm up

**Pick one team member to complete the steps in this section while the others contribute to the discussion but do not actually touch the files on their computer.**

Before we introduce the data, let's warm up with some simple exercises.

### YAML

Open the R Markdown (Rmd) file in your project, change the author name to your **team** name, and knit the document.

### Commiting and pushing changes

- Go to the **Git** pane in your RStudio. 
- View the **Diff** and confirm that you are happy with the changes.
- Add a commit message like "Update team name" in the **Commit message** box and hit **Commit**.
- Click on **Push**. This will prompt a dialogue box where you first need to enter your user name, and then your password.

### Pulling changes

Now, the remaining team members who have not been concurrently making these changes on their projects should click on the **Pull** button in their Git pane and observe that the changes are now reflected on their projects as well.

## Set a seed!

In this lab we'll be generating random samples. The last thing you want is those samples to change every time you knit your document. So, you should set a seed. There's an R chunk in your R Markdown file set aside for this. Locate it and add a seed. Make sure all members in a team are using the same seed so that you don't get merge conflicts and your results match up for the narratives.

## The data

Load the `ncbirths` data from the `openintro` package:

```{r load-data}
data(ncbirths)
```

We have observations on 13 different variables, some categorical and some 
numerical. The meaning of each variable is as follows.

variable         | description
---------------- | ---------------------------------------------
`fage`           | father's age in years.
`mage`           | mother's age in years.
`mature`         | maturity status of mother.
`weeks`          | length of pregnancy in weeks.
`premie`         | whether the birth was classified as premature (premie) or full-term.
`visits`         | number of hospital visits during pregnancy.
`marital`        | whether mother is `married` or `not married` at birth.
`gained`         | weight gained by mother during pregnancy in pounds.
`weight`         | weight of the baby at birth in pounds.
`lowbirthweight` | whether baby was classified as low birthweight (`low`) or not (`not low`).
`gender`         | gender of the baby, `female` or `male`.
`habit`          | status of the mother as a `nonsmoker` or a `smoker`.
`whitemom`       | whether mom is `white` or `not white`.

## Exercises

1.  What are the cases in this data set? How many cases are there in our sample?

The first step in the analysis of a new dataset is getting acquanted with the data.
Make summaries of the variables in your dataset, determine  which variables are 
categorical and which are numerical. For numerical variables, are there outliers? 
If you aren't sure or want to take a closer look at the data, make a graph.

### Baby weights

```{marginfigure}
Wen, Shi Wu, Michael S. Kramer, and Robert H. Usher. "Comparison of birth weight distributions between Chinese and Caucasian infants." American Journal of Epidemiology 141.12 (1995): 1177-1187.
````

A 1995 study suggestes that average weight of Caucasian babies born in the US is 3,369 grams (7.43 pounds). In this dataset we only have information on mother's race, so we will make the simplifying assumption that babies of Caucasian mothers are also Caucasian, i.e. `whitemom = "white"`. 

We want to evaluate whether the average weight of Caucasian babies has changed since 1995.

Our null hypothesis should state "there is nothing going on", i.e. no change since 1995: $H_0: \mu = 7.43~pounds$.

Our alternative hypothesis should reflect the research question, i.e. some change since 1995. Since the research question doesn't state a direction for the change, we use a two sided alternative hypothesis: $H_A: \mu \ne 7.43~pounds$.

3. Create a filtered data frame called `ncbirths_white` that contain data only from white mothers. Then, calculate the mean of the weights of their babies.

4. Are the conditions necessary for conducting simulation based inference satisfied? Explain your reasoning.

Let's discuss how this test would work. Our goal is to simulate a null distribution of sample means that is centered at the null value of 7.43 pounds. In order to do so, we 

- take a bootstrap sample of from the original sample,
- calculate this bootstrap sample's mean,
- repeat these two steps a large number of times to create a bootstrap distribution of means centered at the observed sample mean,
- shift this distribution to be centered at the null value by substracting / adding X to all boostrap mean (X = difference between mean of bootstrap distribution and null value), and
- calculate the p-value as the proportion of bootstrap samples that yielded a sample mean at least as extreme as the observed sample mean.

5. Run the appropriate hypothesis test, visualize the null distribution, calculate the p-value, and interpret the results in context of the data and the hypothesis test.

### Baby weight vs. smoking

Consider the possible relationship between a mother's smoking habit and the 
weight of her baby. Plotting the data is a useful first step because it helps 
us quickly visualize trends, identify strong associations, and develop research
questions.

6.  Make side-by-side boxplots displaying the relationship between `habit` and 
`weight`. What does the plot highlight about the relationship between these 
two variables?

7. Before moving forward, save a version of the dataset omitting observations where there are NAs for `habit`. You can call this version `ncbirths_habitgiven`.

The box plots show how the medians of the two distributions compare, but we can
also compare the means of the distributions using the following to 
first group the data by the `habit` variable, and then calculate the mean
`weight` in these groups using.

```{r habit-means, eval=FALSE}
ncbirths_habitgiven %>%
  group_by(habit) %>%
  summarise(mean_weight = mean(weight))
```

There is an observed difference, but is this difference statistically 
significant? In order to answer this question we will conduct a hypothesis test
.

7. Write the hypotheses for testing if the average weights of babies born to 
smoking and non-smoking mothers are different.

8. Are the conditions necessary for conducting simulation based inference satisfied? Explain your reasoning.

9. Run the appropriate hypothesis test, calculate the p-value, and interpret the results in context of the data and the hypothesis test.

10. Construct a 95% confidence interval for the difference between the average weights of babies born to smoking and non-smoking mothers.

### Baby weight vs. mother's age
 
In this portion of the analysis we focus on two variables. The first one is `maturemom`.

11. First, a non-inference task: Determine the age cutoff for younger and mature 
mothers. Use a method of your choice, and explain how your method works.

The other variable of interest is `lowbirthweight`.

12. Conduct a hypothesis test evaluating whether the proportion of low birth weight babies is higher for mature mothers. State the hypotheses, verify the conditions, run the test and calculate the p-value, and state your conclusion in context of the research question. Use $\alpha = 0.05$. If you find a significant difference, costruct a confidence interval, at the equivalent level to the hypothesis test, for the difference between the proportions of low birth weight babies between mature and younger moms, and interpret this interval in context of the data.
