---
title: "Bechdel"
author: "Mine Çetinkaya-Rundel"
format: html
editor: visual
---

In this mini analysis we work with the data used in the FiveThirtyEight story titled ["The Dollar-And-Cents Case Against Hollywood's Exclusion of Women"](https://fivethirtyeight.com/features/the-dollar-and-cents-case-against-hollywoods-exclusion-of-women/).
Your task is to fill in the blanks denoted by `___`.

## Data and packages

We start with loading the packages we'll use.

```{r}
#| label: load-packages
#| warning: false
#| message: false

library(fivethirtyeight)
library(tidyverse)
```

The dataset contains information on `r nrow(bechdel)` movies released between `r min(bechdel$year)` and `r max(bechdel$year)`.
However we'll focus our analysis on movies released between 1990 and 2013.

```{r}
bechdel90_13 <- bechdel %>% 
  filter(between(year, 1990, 2013))
```

There are `___` such movies.

The financial variables we'll focus on are the following:

-   `budget_2013`: Budget in 2013 inflation adjusted dollars
-   `domgross_2013`: Domestic gross (US) in 2013 inflation adjusted dollars
-   `intgross_2013`: Total International (i.e., worldwide) gross in 2013 inflation adjusted dollars

And we'll also use the `binary` and `clean_test` variables for **grouping**.

## Analysis

Let's take a look at how median budget and gross vary by whether the movie passed the Bechdel test, which is stored in the `binary` variable.

```{r}
bechdel90_13 %>%
  group_by(binary) %>%
  summarise(
    med_budget = median(budget_2013),
    med_domgross = median(domgross_2013, na.rm = TRUE),
    med_intgross = median(intgross_2013, na.rm = TRUE)
    )
```

Next, let's take a look at how median budget and gross vary by a more detailed indicator of the Bechdel test result.
This information is stored in the `clean_test` variable, which takes on the following values:

-   `ok` = passes test
-   `dubious`
-   `men` = women only talk about men
-   `notalk` = women don't talk to each other
-   `nowomen` = fewer than two women

```{r}
bechdel90_13 %>%
  #group_by(___) %>%
  summarise(
    med_budget = median(budget_2013),
    med_domgross = median(domgross_2013, na.rm = TRUE),
    med_intgross = median(intgross_2013, na.rm = TRUE)
    )
```

In order to evaluate how return on investment varies among movies that pass and fail the Bechdel test, we'll first create a new variable called `roi` as the ratio of the gross to budget.

```{r}
bechdel90_13 <- bechdel90_13 %>%
  mutate(roi = (intgross_2013 + domgross_2013) / budget_2013)
```

Let's see which movies have the highest return on investment.

```{r}
bechdel90_13 %>%
  arrange(desc(roi)) %>% 
  select(title, roi, year)
```

Below is a visualization of the return on investment by test result, however it's difficult to see the distributions due to a few extreme observations.

```{r}
#| warning: false

ggplot(data = bechdel90_13, 
       mapping = aes(x = clean_test, y = roi, color = binary)) +
  geom_boxplot() +
  labs(
    title = "Return on investment vs. Bechdel test result",
    x = "Detailed Bechdel result",
    y = "___",
    color = "Binary Bechdel result"
    )
```

What are those movies with *very* high returns on investment?

```{r}
bechdel90_13 %>%
  filter(roi > 400) %>%
  select(title, budget_2013, domgross_2013, year)
```

Zooming in on the movies with `roi < ___` provides a better view of how the medians across the categories compare:

```{r}
#| warning: false

ggplot(data = bechdel90_13, mapping = aes(x = clean_test, y = roi, color = binary)) +
  geom_boxplot() +
  labs(
    title = "Return on investment vs. Bechdel test result",
    subtitle = "___", # Something about zooming in to a certain level
    x = "Detailed Bechdel result",
    y = "Return on investment",
    color = "Binary Bechdel result"
    ) +
  coord_cartesian(ylim = c(0, 15))
```
