---
title: "Favourite foods"
author: "Mine Çetinkaya-Rundel"
date: "`r Sys.Date()`"
output: 
  html_document: 
    toc: yes
    toc_float: yes
---

```{r load-packages, message=FALSE}
library(tidyverse)
library(readxl)
```

## Exercise 1

* Read in the Excel file called `favourite-food.xlsx` from the `data-raw/` folder.

```{r label-me1, eval=FALSE}
fav_food <- read_excel(___)
fav_food
```

* Clean up `NA`s and make sure you're happy with variable types.

```{r label-me2, eval=FALSE}
fav_food <- read_excel(___, ___)
fav_food 
```

* Convert SES (socio economic status) to a factor variables with levels in the 
following order: `Low`, `Middle`, `High`.

```{r label-me3}
# add code here
```

* Write out the resulting data frame to `favourite-food.csv` in the `data/` folder.

```{r label-me4}
# add code here
```

* Finally, read `favourite-food.csv` back in from the `data/` folder and 
observe the variable types. Are they as you left them?

```{r label-me5}
# add code here
```

## Exercise 2

* Read in the Excel file called `favourite-food.xlsx` from the `data-raw/` folder, 
taking care of `NA`s and variable types.

```{r label-me6, eval=FALSE}
fav_food <- read_excel(___, ___)
fav_food 
```

* Convert SES (socio economic status) to a factor variables with levels in the 
following order: `Low`, `Middle`, `High`.

```{r label-me7}
# add code here
```

* Write out the resulting data frame to `favourite-food.rds` in the `data/` folder.

```{r label-me8}
# add code here
```

* Read `favourite-food.rds` back in from the `data/` folder and observe the 
variable types. Are they as you left them?

```{r label-me9}
# add code here
```
