---
title: "Simulation based inference review <br> `r emo::ji('books')`"
author: ""
output:
  xaringan::moon_reader:
    css: "../slides.css"
    lib_dir: libs
    nature:
      ratio: "16:9"
      highlightLines: true
      highlightStyle: solarized-light
      countIncrementalSlides: false
---

```{r child = "../setup.Rmd"}
```

```{r packages, echo=FALSE, message=FALSE, warning=FALSE}
library(tidyverse)
library(broom)
library(knitr)
library(DT)
library(emo)
library(openintro)
library(infer)
```

class: middle

# Simulation based inference review

---

## What do you want to do?

- Estimation -> Confidence interval

- Decision -> Hypothesis test

- First step: Ask the following questions

  1. How many variables?
  2. What type(s) of variable(s)?
  3. What is the research question?

---

## Data: NC births

The dataset is in the `openintro` package.

```{r}
glimpse(ncbirths)
```

---

## Length of gestation

```{r echo=FALSE, warning=FALSE}
ggplot(data = ncbirths, aes(x = weeks)) +
  geom_histogram(binwidth = 1)
```

```{r echo=FALSE}
ncbirths %>%
  filter(!is.na(weeks)) %>%
  summarise(
    min = min(weeks),
    xbar = round(mean(weeks), 2),
    med = median(weeks),
    s = round(sd(weeks), 2),
    q1 = quantile(weeks, 0.25),
    q3 = quantile(weeks, 0.75),
    max = max(weeks)
  )
```


---

## Length of gestation

.question[
Assuming that this sample is representative of all births in NC, we are 95% confident that the average length of gestation for babies in NC is between ---- and ---- weeks.
]

--

**(1) How many variables?**

--

1 variable: length of gestation, `weeks`

--

**(2) What type(s) of variable(s)?**

--

Numerical

--

**(3) What is the research question?**

--

Estimate the average length of gestation $\rightarrow$ confidence interval

---

## Simulation for CI for a mean

**Goal:** Use bootstrapping to estimate the sampling variability of the mean, i.e. the variability of means taken from the same population with the same sample size.

--

1. Take a bootstrap sample - a random sample taken with replacement from the 
original sample, of the same size as the original sample.

2. Calculate the mean of the bootstrap sample.

3. Repeat steps (1) and (2) many times to create a bootstrap distribution - 
a distribution of bootstrap means.

4. Calculate the bounds of the 95% confidence interval as the middle 95% 
of the bootstrap distribution.

---

## Set a seed first

From the documentation of `set.seed`:

- `set.seed` uses a single integer argument to set as many seeds as are required. There is no guarantee that different values of seed will seed the RNG differently, although any exceptions would be extremely rare.

- Initially, there is no seed; a new one is created from the current time and the process ID when one is required. Hence different sessions will give different simulation results, by default.

```{r}
set.seed(20180326)
```

---

## Computation for CI for a mean

```{r}
boot_means <- ncbirths %>%
  filter(!is.na(weeks)) %>% # remove NAs
  specify(response = weeks) %>%
  generate(reps = 1000, type = "bootstrap") %>%
  calculate(stat = "mean")
ggplot(data = boot_means, aes(x = stat)) +
  geom_histogram(binwidth = 0.03)
```

---

## Length of gestation

```{r}
boot_means %>%
  summarise(
    lower = quantile(stat, 0.025),
    upper = quantile(stat, 0.975)
  )
```

--

Assuming that this sample is representative of all births in NC, we are 95% confident that the average length of gestation for babies in NC is between 38.1 and 38.5 weeks.

---

## Length of gestation, revisited

.question[
The average length of human gestation is 280 days, or 40 weeks, from the first day of the woman's last menstrual period. Do these data provide convincing evidence that average length of gestation for women in NC is different than 40 weeks? Use a significance level of 5%.
]

--

$H_0: \mu = 40$  
$H_A: \mu \ne 40$

--

- We just said, "we are 95% confident that the average length of gestation for babies in NC is between 38.1 and 38.5 weeks".

- Since the null value is outside the CI, we would reject the null hypothesis in favor of the alternative.

- But an alternative, more direct, way of answering this question is using a hypothesis test.

---

## Simulation for HT for a mean

**Goal:** Use bootstrapping to generate a sampling distribution under the assumption of the null hypothesis being true. Then, calculate the p-value to make a decision on the hypotheses.

--

1. Take a bootstrap sample - a random sample taken with replacement from the 
original sample, of the same size as the original sample.

2. Calculate the mean of the bootstrap sample.

3. Repeat steps (1) and (2) many times to create a bootstrap distribution - 
a distribution of bootstrap means.

4. Shift the bootstrap distribution to be centered at the null value by subtracting/adding the difference between the center of the bootstrap distribution and the null value to each bootstrap mean.

5. Calculate the p-value as the proportion of simulations that yield a sample mean at least as extreme as the observed sample mean.

---

## Computation for HT for a mean

```{r fig.height=2}
boot_means_shifted <- ncbirths %>%
  filter(!is.na(weeks)) %>% # remove NAs
  specify(response = weeks) %>%
  hypothesize(null = "point", mu = 40) %>% # hypothesize step
  generate(reps = 1000, type = "bootstrap") %>%
  calculate(stat = "mean")
ggplot(data = boot_means_shifted, aes(x = stat)) +
  geom_histogram(binwidth = 0.03) +
  geom_vline(xintercept = 38.33, color = "red") +
  geom_vline(xintercept = 40 + (40 - 38.33), color = "red")
```

---

## Length of gestation

```{r}
boot_means_shifted %>%
  filter(stat <= 38.33) %>%
  summarise(p_value = 2 * (n() / 1000))
```

--

Since p-value less than the significance level, we reject the null hypothesis. The data provide convincing evidence that the average length of gestation of births in NC is different than 40.

---

## Exercises

Go to RStudio Cloud, make a copy of **NC Births**, and answer the following questions.

1. Do these data provide convincing evidence of a difference in length in gestation between mature and younger moms? Use a significance level of 10%.

2. Estimate the difference in average lengths of gestation between mature and younger moms. Use a significance level equivalent to the hypothesis test above.

3. Do the results of the hypothesis test agree with the result of the confidence interval?

---

## `infer` structure

```{r eval=FALSE}
df %>%
  specify(response, explanatory) %>% # explanatory optional
  generate(reps, type) %>% # type: bootstrap, simulate, or permute
  calculate(stat)
```

- Always start with data frame
- Result is always a data frame with a variable called `stat`
   - See the documentation for `calculate` to see which `stat`istics can be calculated
- For hypothesis testing add a `hypothesize()` step between `specify()` and `generate()`
    - `null = "point"`, and then specify the null value
    - `null = "independence"`
