---
title: "Simpson's paradox"
subtitle: "<br><br> Data Science in a Box"
author: "[datasciencebox.org](https://datasciencebox.org/)"
output:
  xaringan::moon_reader:
    css: ["../xaringan-themer.css", "../slides.css"]
    lib_dir: libs
    anchor_sections: FALSE
    nature:
      ratio: "16:9"
      highlightLines: true
      highlightStyle: solarized-light
      countIncrementalSlides: false
---

```{r child = "../setup.Rmd"}
```

```{r packages, echo = FALSE, message=FALSE, warning=FALSE}
library(tidyverse)
library(scales)
library(kableExtra)
library(dsbox)
```

class: middle

# Case study: Berkeley admission data

---

## Berkeley admission data

- Study carried out by the Graduate Division of the University of California, Berkeley in the early 70’s to evaluate whether there was a gender bias in graduate admissions.
- The data come from six departments. For confidentiality we'll call them A-F. 
- We have information on whether the applicant was male or female and whether they were admitted or rejected. 
- First, we will evaluate whether the percentage of males admitted is indeed higher than females, overall. Next, we will calculate the same percentage for each department.

---

## Data

.pull-left[
```{r message=FALSE, echo=FALSE}
ucbadmit %>%
  print(n = 15)
```
]
.pull-right[
```{r message=FALSE, echo=FALSE}
ucbadmit %>%
  count(gender)

ucbadmit %>%
  count(dept)

ucbadmit %>%
  count(admit)
```
]

---

.question[
What can you say about the overall gender distribution? Hint: Calculate the following probabilities: $P(Admit | Male)$ and $P(Admit | Female)$.
]

```{r}
ucbadmit %>%
  count(gender, admit)
```

---

```{r ucbadmit-overall-prop}
ucbadmit %>%
  count(gender, admit) %>%
  group_by(gender) %>%
  mutate(prop_admit = n / sum(n))
```

- $P(Admit | Female)$ = 0.304
- $P(Admit | Male)$ = 0.445

---

## Overall gender distribution

.panelset[

.panel[.panel-name[Plot]
```{r ref.label="gender-admit", echo = FALSE}
```
]

.panel[.panel-name[Code]
```{r gender-admit, fig.show = "hide"}
ggplot(ucbadmit, aes(y = gender, fill = admit)) +
  geom_bar(position = "fill") + 
  labs(title = "Admit by gender",
       y = NULL, x = NULL)
```
]

]

---

.question[
What can you say about the gender distribution by department ?
]

```{r}
ucbadmit %>%
  count(dept, gender, admit)
```

---

.question[
Let's try again... What can you say about the gender distribution by department?
]

```{r}
ucbadmit %>%
  count(dept, gender, admit) %>%
  pivot_wider(names_from = dept, values_from = n)
```

---

## Gender distribution, by department

.panelset[

.panel[.panel-name[Plot]
```{r ref.label="gender-dept-admit", echo = FALSE}
```
]

.panel[.panel-name[Code]
```{r gender-dept-admit, fig.show = "hide"}
ggplot(ucbadmit, aes(y = gender, fill = admit)) +
  geom_bar(position = "fill") +
  facet_wrap(. ~ dept) +
  scale_x_continuous(labels = label_percent()) +
  labs(title = "Admissions by gender and department",
       x = NULL, y = NULL, fill = NULL) +
  theme(legend.position = "bottom")
```
]

]

---

## Case for gender discrimination?

.pull-left[
```{r ref.label="gender-admit", echo=FALSE, out.width="100%"}
```
]
.pull-right[
```{r ref.label="gender-dept-admit", echo=FALSE, out.width="100%"}
```
]

---

## Closer look at departments

.panelset[

.panel[.panel-name[Output]
```{r ref.label="gender-dept-admit-table", echo = FALSE}
```
]

.panel[.panel-name[Code]
```{r gender-dept-admit-table, eval = FALSE}
ucbadmit %>%
  count(dept, gender, admit) %>%
  group_by(dept, gender) %>%
  mutate(
    n_applied  = sum(n),
    prop_admit = n / n_applied
    ) %>%
  filter(admit == "Admitted") %>%
  rename(n_admitted = n) %>%
  select(-admit) %>%
  print(n = 12)
```
]

]

---

class: middle

# Simpson's paradox

---

## Relationship between two variables

.pull-left[
```{r echo=FALSE, message=FALSE}
df <- tribble(
  ~x, ~y, ~z,
  2,   4,  "A",
  3,   3,  "A",
  4,   2,  "A",
  5,   1,  "A",
  6,   11, "B",
  7,   10, "B",
  8,   9,  "B",
  9,   8,  "B"
)
df
```
]
.pull-right[
```{r echo=FALSE, out.width="100%"}
ggplot(df) +
  geom_point(aes(x = x, y = y), color = "darkgray", size = 5) +
  theme_minimal()
```
]

---

## Relationship between two variables

.pull-left[
```{r echo=FALSE, message=FALSE}
df
```
]
.pull-right[
```{r echo=FALSE, out.width="100%"}
ggplot(data = df) +
  geom_point(aes(x = x, y = y), color = "darkgray", size = 5) +
  geom_smooth(aes(x = x, y = x), color = "darkgray") +
  theme_minimal()
```
]

---

## Considering a third variable

.pull-left[
```{r echo=FALSE, message=FALSE}
df
```
]
.pull-right[
```{r echo=FALSE, out.width="100%"}
ggplot(data = df) +
  geom_point(aes(x = x, y = y, color = z), size = 5) +
  geom_smooth(aes(x = x, y = x), method = "lm", color = "darkgray") +
  theme_minimal()
```
]

---

## Relationship between three variables

.pull-left[
```{r echo=FALSE, message=FALSE}
df
```
]
.pull-right[
```{r echo=FALSE, out.width="100%"}
ggplot(data = df) +
  geom_point(aes(x = x, y = y, color = z), size = 5) +
  geom_smooth(aes(x = x, y = x), method = "lm", color = "darkgray") +
  geom_smooth(aes(x = x, y = y, color = z), method = "lm") +
  theme_minimal()
```
]

---

## Simpson's paradox

- Not considering an important variable when studying a relationship can result in **Simpson's paradox**
- Simpson's paradox illustrates the effect that omission of an explanatory variable can have on the measure of association between another explanatory variable and a response variable
- The inclusion of a third variable in the analysis can change the apparent relationship between the other two variables

---

class: middle

# Aside: `group_by()` and `count()`

---

## What does group_by() do?

`group_by()` takes an existing data frame and converts it into a grouped data frame where subsequent operations are performed "once per group"

.pull-left[
```{r}
ucbadmit
```
]
.pull-right[
```{r}
ucbadmit %>% 
  group_by(gender)
```
]

---

## What does group_by() not do?

`group_by()` does not sort the data, `arrange()` does

.pull-left[
```{r}
ucbadmit %>% 
  group_by(gender)
```
]
.pull-right[
```{r}
ucbadmit %>% 
  arrange(gender)
```
]

---

## What does group_by() not do?

`group_by()` does not create frequency tables, `count()` does

.pull-left[
```{r}
ucbadmit %>% 
  group_by(gender)
```
]
.pull-right[
```{r}
ucbadmit %>% 
  count(gender)
```
]

---

## Undo grouping with ungroup()

.pull-left[
```{r}
ucbadmit %>%
  count(gender, admit) %>%
  group_by(gender) %>%
  mutate(prop_admit = n / sum(n)) %>%
  select(gender, prop_admit)
```
]
.pull-right[
```{r}
ucbadmit %>%
  count(gender, admit) %>%
  group_by(gender) %>%
  mutate(prop_admit = n / sum(n)) %>%
  select(gender, prop_admit) %>%
  ungroup()
```
]

---

## count() is a short-hand

`count()` is a short-hand for `group_by()` and then `summarise()` to count the number of observations in each group

.pull-left[
```{r}
ucbadmit %>%
  group_by(gender) %>%
  summarise(n = n()) 
```
]
.pull-right[
```{r}
ucbadmit %>%
  count(gender)
```
]

---

## count can take multiple arguments

.pull-left[
```{r}
ucbadmit %>%
  group_by(gender, admit) %>%
  summarise(n = n()) 
```
]
.pull-right[
```{r}
ucbadmit %>%
  count(gender, admit)
```
]

---

## `summarise()` after `group_by()`

- `count()` ungroups after itself
- `summarise()` peels off one layer of grouping by default, or you can specify a different behaviour


```{r message=TRUE}
ucbadmit %>%
  group_by(gender, admit) %>%
  summarise(n = n()) 
```
