---
title: "Tidytext analysis"
author: "Dr. Çetinkaya-Rundel"
output:
  xaringan::moon_reader:
    css: "../slides.css"
    logo: "../logo/dsbox-logo.png"
    lib_dir: libs
    nature:
      highlightStyle: github
      highlightLines: true
      countIncrementalSlides: false
---

```{r packages, echo=FALSE, message=FALSE, warning=FALSE}
library(tidyverse)
library(broom)
library(knitr)
library(DT)
library(emo)
library(openintro)
library(infer)
library(gridExtra)
```

```{r setup, include=FALSE}
# R options
options(
  htmltools.dir.version = FALSE, # for blogdown
  show.signif.stars = FALSE,     # for regression output
  warm = 1
  )
# Set dpi and height for images
opts_chunk$set(fig.height = 2.5, fig.width = 5, dpi = 300) 
# ggplot2 color palette with gray
color_palette <- list(gray = "#999999", 
                      salmon = "#E69F00", 
                      lightblue = "#56B4E9", 
                      green = "#009E73", 
                      yellow = "#F0E442", 
                      darkblue = "#0072B2", 
                      red = "#D55E00", 
                      purple = "#CC79A7")
htmltools::tagList(rmarkdown::html_dependency_font_awesome())
# For magick
dev.off <- function(){
  invisible(grDevices::dev.off())
}
# For ggplot2
ggplot2::theme_set(ggplot2::theme_bw())
```

## Announcements

- Office hours during finals week:
  - TAs have their regular office hours listed on the course website
  - I will have OH 12-1:30pm Wednesday and Friday

- All grades (except final project) should be in by the end of the day on Thursday May 3
  - At that point check Sakai to make sure your grades are correctly recorded
  - If you catch any issues (in recorded grade) Slack me/TAs -- for regrade
  requests use the usual regrade process, this is just for errors/missingness
  in recorded grades
  
- Extra credit 2: If we can get *both* course and TA evaluation response rates 
to above 80% (currently we’re at 40% for the course evals) everyone gets +5 
pts on their total (not average) HW score. (We're at 45% as of 11:30am today)

---

## Announcements (cont.)

- Presentation schedule: Saturday, May 5
  - Section 1 teams: 9am - 10am
  - Section 2 teams: 10am - 11am
  - Section 3 teams: 11am - 12pm
  
- There will be one more peer eval, specificall for the project, due the evening
of May 5

---

class: center, middle

# Tidytext analysis

---

## Packages

In addition to `tidyverse` we will be using three other packages today

```{r}
library(tidytext)
library(geniusR) # https://github.com/JosiahParry/geniusR
library(gutenbergr)
```

---

## Tidytext

- Using tidy data principles can make many text mining tasks easier, more effective, and consistent with tools already in wide use.

- Learn more at https://www.tidytextmining.com/.

---

## What is tidy text?

```{r}
text <- c("Take me out tonight",
          "Where there's music and there's people",
          "And they're young and alive",
          "Driving in your car",
          "I never never want to go home",
          "Because I haven't got one",
          "Anymore")

text
```

---

## What is tidy text?

```{r}
text_df <- tibble(line = 1:7, text = text)

text_df
```

---

## What is tidy text?

```{r}
text_df %>%
  unnest_tokens(word, text)
```

---

class: center, middle

# Analyzing lyrics of one artist

---

## Let's get more data

We'll use the `geniusR` package to get song lyric data from [Genius](https://genius.com/).

- `genius_album()` allows you to download the lyrics for an entire album in a 
tidy format. 

- Input: Two arguments artists and album. Supply the quoted name of artist 
and the album (if it gives you issues check that you have the album name and 
artists as specified on [Genius](https://genius.com/)).

- Output: A tidy data frame with three columns:
    - `title`: track name
    - `track_n`: track number
    - `text`: lyrics

---

## I have no idea who this person is!

```{r cache=TRUE}
marc <- genius_album(
  artist = "Marc E. Bassy", 
  album = "Gossip Columns"
  )
marc
```

---

## Save for later

```{r}
marc <- marc %>%
  mutate(
    album = "Gossip Columns",
    artist = "Marc E. Bassy"
    )
```

---

## What songs are in the album?

```{r}
marc %>%
  distinct(track_title)
```

---

## How long are Marc's songs?

Length measured by number of lines

```{r}
marc %>%
  count(track_title) %>%
  arrange(desc(n))
```

---

## Tidy up your lyrics!

```{r}
marc_lyrics <- marc %>%
  unnest_tokens(word, lyric)

marc_lyrics
```

---

## What are the most common words?

```{r}
marc_lyrics %>%
  count(word) %>%
  arrange(desc(n))
```

---

## Stop words

- In computing, stop words are words which are filtered out before or after processing of natural language data (text).

- They usually refer to the most common words in a language, but there is not a single list of stop words used by all natural language processing tools.

---

## Spanish stop words

```{r}
get_stopwords(language = "es")
```

---

## Various lexicons

See `?get_stopwords` for more info.

```{r}
get_stopwords(source = "smart")
```

---

## What are the most common words?

```{r}
marc_lyrics %>%
  anti_join(get_stopwords(source = "smart")) %>%
  count(word) %>%
  arrange(desc(n))
```

---

## What are the most common words?

```{r eval=FALSE}
marc_lyrics %>%
  anti_join(get_stopwords(source = "smart")) %>%
  count(word) %>%
  arrange(desc(n)) %>%
  top_n(20) %>%
  ggplot(aes(fct_reorder(word, n), n)) +
    geom_col() +
    coord_flip() + 
    theme_minimal() +
    labs(title = "Frequency of Marc E. Bassy's lyrics",
         subtitle = "`Love` tops the chart",
         y = "",
         x = "")
```

---

```{r echo=FALSE, message=FALSE, fig.height=4.5}
marc_lyrics %>%
  anti_join(get_stopwords(source = "smart")) %>%
  count(word) %>%
  arrange(desc(n)) %>%
  top_n(20) %>%
  ggplot(aes(fct_reorder(word, n), n)) +
    geom_col() +
    coord_flip() + 
    theme_minimal() +
    labs(title = "Frequency of Marc E. Bassy's lyrics",
         subtitle = "`Love` tops the chart",
         y = "",
         x = "")
```

---

## Sentiment analysis

- One way to analyze the sentiment of a text is to consider the text as a combination of its individual words 

- and the sentiment content of the whole text as the sum of the sentiment content of the individual words

---

## Sentiment lexicons

.pull-left[
```{r}
get_sentiments("afinn")
```
]
.pull-right[
```{r}
get_sentiments("bing") 
```
]

---

## Sentiment lexicons

.pull-left[
```{r}
get_sentiments("nrc")
```
]
.pull-right[
```{r}
get_sentiments("loughran") 
```
]

---

## Sentiments in Marc's lyrics

```{r}
marc_lyrics %>%
  inner_join(get_sentiments("bing")) %>%
  count(sentiment, word) %>%
  arrange(desc(n))
```

---

## Visualizing sentiments

```{r echo=FALSE, message=FALSE}
marc_lyrics %>%
  inner_join(get_sentiments("bing")) %>%
  count(sentiment, word) %>%
  arrange(desc(n)) %>%
  group_by(sentiment) %>%
  top_n(10) %>%
  ungroup() %>%
  ggplot(aes(fct_reorder(word, n), n, fill = sentiment)) +
    geom_col() +
    coord_flip() +
    facet_wrap(~ sentiment, scales = "free") +
    theme_minimal() +
    labs(title = "Sentiments in Marc E. Bassy's lyrics",
         x = "")
```

---

## Visualizing sentiments

```{r eval=FALSE}
marc_lyrics %>%
  inner_join(get_sentiments("bing")) %>%
  count(sentiment, word) %>%
  arrange(desc(n)) %>%
  group_by(sentiment) %>%
  top_n(10) %>%
  ungroup() %>%
  ggplot(aes(fct_reorder(word, n), n, fill = sentiment)) +
    geom_col() +
    coord_flip() +
    facet_wrap(~ sentiment, scales = "free") +
    theme_minimal() +
    labs(title = "Sentiments in Marc E. Bassy's lyrics",
         x = "")
```

---

class: center, middle

# Comparing lyrics across artists

---

## Get more data

Get data from two more artists:

```{r cache=TRUE}
quinn <- genius_album(artist = "Quinn XCII", album = "The Story of Us") %>%
  mutate(artist = "Quinn XCII", album = "The Story of Us")
seeb <- genius_album(artist = "Seeb", album = "Nice to Meet You EP") %>%
  mutate(artist = "Seeb", album = "Nice to Meet You EP")
```

--

Combine data:

```{r}
ldoc <- bind_rows(marc, quinn, seeb)
```

---

## LDOC lyrics

```{r}
ldoc_lyrics <- ldoc %>%
  unnest_tokens(word, lyric)
```

---

## Common LDOC lyrics

Without stop words:

```{r}
ldoc_lyrics %>%
  anti_join(get_stopwords(source = "smart")) %>%
  count(artist, word, sort = TRUE) # alternative way to sort
```

---

## Common LDOC lyrics

With stop words:

```{r}
ldoc_lyrics_counts <- ldoc_lyrics %>%
  count(artist, word, sort = TRUE)
```

---

## What is a document about?

- Term frequency
- Inverse document frequency

$$idf(\text{term}) = \ln{\left(\frac{n_{\text{documents}}}{n_{\text{documents containing term}}}\right)}$$

tf-idf is about comparing **documents** within a **collection**.

---

## Calculating tf-idf

This is not that exciting... What's the issue?

```{r}
ldoc_words <- ldoc_lyrics_counts %>%
  bind_tf_idf(word, artist, n)

ldoc_words
```

---

## Re-calculating tf-idf

```{r}
ldoc_words %>%
  bind_tf_idf(word, artist, n) %>%
  arrange(-tf_idf)
```

---

## 

```{r echo=FALSE,message=FALSE, fig.height=4.5}
ldoc_words %>% 
  group_by(artist) %>% 
  top_n(10) %>% 
  ungroup() %>% 
  ggplot(aes(fct_reorder(word, tf_idf), tf_idf, fill = artist)) + 
    geom_col(alpha = 0.9, show.legend = FALSE) + coord_flip() + 
    facet_wrap(~artist, ncol=1, scales = "free") + 
    scale_y_continuous(expand = c(0,0)) + 
    theme_minimal() +
    labs(x = NULL, y = "tf-idf") +
    ylim(c(0, 0.017))
```

---

## Acknowledgements

- Julia Silge: https://github.com/juliasilge/tidytext-tutorial

- Julia Silge and David Robinson: https://www.tidytextmining.com/

- Josiah Parry: https://github.com/JosiahParry/geniusR





