---
title: "Multiple linear regression + Model selection"
output:
  xaringan::moon_reader:
    css: "../slides.css"
    logo: img/sta199-logo-hex.png
    lib_dir: libs
    nature:
      highlightStyle: github
      highlightLines: true
      countIncrementalSlides: false
---

```{r packages, echo=FALSE, message=FALSE, warning=FALSE}
library(tidyverse)
library(broom)
```

```{r setup, include=FALSE}
# R options
options(
  htmltools.dir.version = FALSE, # for blogdown
  show.signif.stars = FALSE,     # for regression output
  warn = 1
  )
# Set dpi and height for images
knitr::opts_chunk$set(fig.height = 2.5, fig.width = 5, dpi = 300) 
# ggplot2 color palette with gray
color_palette <- list(gray = "#999999", 
                      salmon = "#E69F00", 
                      lightblue = "#56B4E9", 
                      green = "#009E73", 
                      yellow = "#F0E442", 
                      darkblue = "#0072B2", 
                      red = "#D55E00", 
                      purple = "#CC79A7")
htmltools::tagList(rmarkdown::html_dependency_font_awesome())
# For magick
dev.off <- function(){
  invisible(grDevices::dev.off())
}
# For ggplot2
ggplot2::theme_set(ggplot2::theme_bw())
```

class: center, middle

## Data cleaning

---

## Load data

```{r message=FALSE}
pp <- read_csv("../../data/paris-paintings.csv", 
               na = c("n/a", "", "NA"))
```

---

## Shape and material

Collapse levels of `Shape` and `mat`erial variables with `forcats::fct_collapse`:

.small[
```{r}
pp <- pp %>%
  mutate(
    Shape = fct_collapse(Shape, oval = c("oval", "ovale"),
                                round = c("round", "ronde"),
                                squ_rect = "squ_rect",
                                other = c("octogon", "octagon", "miniature")),
    mat = fct_collapse(mat, metal = c("a", "br", "c"),
                            canvas = c("co", "t", "ta"),
                            paper = c("p", "ca"),
                            wood = "b",
                            other = c("e", "g", "h", "mi", "o", "pa", "v", "al", "ar", "m"))
  )
```
]

---

## Review fixes

.pull-left[
```{r}
pp %>%
  count(Shape)
```
]

.pull-right[
```{r}
pp %>%
  count(mat)
```
]

---

class: center, middle

## Review

---

## Main effects, numerical predictors

```{r}
m_main_n <- lm(log(price) ~ Width_in + Height_in, data = pp)
tidy(m_main_n) %>%
  select(term, estimate) %>%
  mutate(estimate = round(estimate, 3))
```

---

## Visualizing the model

```{r echo=FALSE}
library(plotly)
library(widgetframe)
```

```{r echo=FALSE}
p <- plot_ly(pp, x = ~Width_in, y = ~Height_in, z = ~log(price),
        marker = list(size = 3,
                       color = "lightgray",
                       alpha = 0.5,
                       line = list(color = "gray",
                                   width = 2))) %>%
  add_markers() %>%
  layout(scene = list(xaxis = list(title = "Width (in)"),
                      yaxis = list(title = "Height (in)"),
                      zaxis = list(title = "Log(price)"))) %>%
  config(displayModeBar = FALSE)

frameWidget(p, width = '100%', height = '70%')
```


---

## Main effects, numerical and categorical predictors

.small[
```{r}
pp_Surf_lt_5000 <- pp %>% filter(Surface < 5000)
m_main <- lm(log(price) ~ Surface + factor(artistliving), data = pp_Surf_lt_5000)
tidy(m_main) %>%
  select(term, estimate) %>%
  mutate(exp_estimate = round(exp(estimate), 4))
```

```{r include=FALSE}
m_main_coefs <- tidy(m_main) %>% 
  mutate(exp_estimate = round(exp(estimate), 4)) %>%
  select(term, exp_estimate)
```

- All else held constant, for each additional square inch in painting's surface area, the price of the painting is predicted, on average, to be higher by a factor of `r m_main_coefs %>% filter(term == "Surface") %>% pull(exp_estimate)`.

- All else held constant, paintings by a living artist are predicted, on average, to be higher by a factor of `r m_main_coefs %>% filter(term == "factor(artistliving)1") %>% pull(exp_estimate)` compared to paintings by an artist who is no longer alive.

- Paintings that are by an artist who is not alive and that have a surface area of 0 square inches are predicted, on average, to be `r m_main_coefs %>% filter(term == "(Intercept)") %>% pull(exp_estimate)` livres.
]

---

## What went wrong?

.question[
Why is our linear regression model different from what we got from `geom_smooth(method = "lm")`?
]

.pull-left[
```{r echo=FALSE, fig.height=4}
ggplot(pp_Surf_lt_5000, aes(x = Surface, y = log(price), color = factor(artistliving))) + 
  geom_point(alpha = 0.3) +
  geom_smooth(method = "lm") +
  labs(x = "Surface", y = "Log(price)", color = "Living artist")
```
]
.pull-right[
```{r echo=FALSE, fig.height=4}
m_pr <- lm(log(price) ~ Surface + factor(artistliving), data = pp_Surf_lt_5000)
m_pr_aug <- augment(m_pr)

ggplot(data = m_pr_aug, mapping = aes(y = log.price., x = Surface, color = factor.artistliving.)) +
  geom_point(alpha = 0.3) +
  geom_line(aes(y = .fitted))
```
]

---

## What went wrong? (cont.)

- The way we specified our model only lets `artistliving` affect the intercept.

- Model implicitly assumes that paintings with living and deceased artists have the *same slope* and only allows for *different intercepts*.  

- What seems more appropriate in this case? 
    
    * Same slope and same intercept for both colors
    
    * Same slope and different intercept for both colors
    
    * Different slope and different intercept for both colors?

---

## Interacting explanatory variables

- Including an interaction effect in the model allows for different slopes, i.e. 
nonparallel lines.

- This implies that the regression coefficient for an explanatory variable would 
change as another explanatory variable changes.

- This can be accomplished by adding an interaction variable: the product of two 
explanatory variables.

---

## Interaction: surface * artist living

.small[
```{r fig.height=2.5}
ggplot(data = pp_Surf_lt_5000,
       mapping = aes(y = log(price), x = Surface, 
                     color = factor(artistliving))) +
  geom_point(alpha = 0.3) +
  geom_smooth(method = "lm") +
  labs(x = "Surface", y = "Log(price)", color = "Living artist")
```
]

---

## Modeling with interaction effects

.small[
```{r}
m_int <- lm(log(price) ~ Surface + factor(artistliving) +
              Surface * factor(artistliving), 
            data = pp_Surf_lt_5000)
tidy(m_int) %>% 
  select(term, estimate)
```
]

$$ \widehat{log(price)} = 4.91 + 0.00021~surface - 0.126~artistliving $$
$$+ ~ 0.00048~surface \times artistliving $$

---

## Interpretation of interaction effects

- Rate of change in price as the surface area of the painting increases does 
vary between paintings by living and non-living artists (different slopes), 

- Some paintings by living artists are more expensive than paintings by
non-living artists, and some are not (different intercept).

.small[
.pull-left[
- Non-living artist: 
$\widehat{log(price)} = 4.91 + 0.00021~surface$
$- 0.126 \times 0 + 0.00048~surface \times 0$
$= 4.91 + 0.00021~surface$

- Living artist: 
$\widehat{log(price)} = 4.91 + 0.00021~surface$
$- 0.126 \times 1 + 0.00048~surface \times 1$
$= 4.91 + 0.00021~surface$
$- 0.126 + 0.00048~surface$
$= 4.784 + 0.00069~surface$
]
.pull-right[
```{r fig.height=4, echo = FALSE}
ggplot(data = pp_Surf_lt_5000,
       aes(y = log(price), x = Surface, color = factor(artistliving))) +
  geom_point(alpha = 0.3) +
  stat_smooth(method = "lm", fullrange = TRUE)
```
]
]

---

## Third order interactions

- Can you? Yes

- Should you? Probably not if you want to interpret these interactions in context
of the data.

---

class: center, middle

# Quality of fit in MLR

---

## $R^2$

- $R^2$ is the percentage of variability in the response variable explained by the 
regression model.

```{r}
glance(m_main)$r.squared
glance(m_int)$r.squared
```

--

- Clearly the model with interactions has a higher $R^2$.

--

- However using $R^2$ for model selection in models with multiple explanatory 
variables is not a good idea as $R^2$ increases when **any** variable is added to the 
model.

---

## $R^2$ - first principles

$$ R^2 =  \frac{ SS\_{Reg} }{ SS\_{Total} } = 1 - \left( \frac{ SS\_{Error} }{ SS\_{Total} } \right) $$

.question[
Calculate $R^2$ based on the output below.
]

```{r digits = 3}
anova(m_main)
```

---

## Adjusted $R^2$

$$ R^2\_{adj} = 1 - \left( \frac{ SS\_{Error} }{ SS\_{Total} } \times \frac{n - 1}{n - k - 1} \right), $$

where $n$ is the number of cases and $k$ is the number of predictors in the model

--

- Adjusted $R^2$ doesn't increase if the new variable does not provide any new 
informaton or is completely unrelated.

--

- This makes adjusted $R^2$ a preferable metric for model selection in multiple
regression models.

---

## In pursuit of Occam's Razor

- Occam's Razor states that among competing hypotheses that predict equally well, 
the one with the fewest assumptions should be selected.

- Model selection follows this principle.

- We only want to add another variable to the model if the addition of that
variable brings something valuable in terms of predictive power to the model.

- In other words, we prefer the simplest best model, i.e. **parsimonious** model.

---

## Comparing models

It appears that adding the interaction actually increased adjusted $R^2$, so we 
should indeed use the model with the interactions.

```{r}
glance(m_main)$adj.r.squared
glance(m_int)$adj.r.squared
```

---

class: center, middle

# Model selection

---

## Backwards elimination

- Start with **full** model (including all candidate explanatory variables and all
candidate interactions)

- Remove one variable at a time, and select the model with the highest adjusted $R^2$

- Continue until adjusted $R^2$ does not increase

---

## Forward selection

- Start with **empty** model

- Add one variable (or interaction effect) at a time, and select the model with the 
highest adjusted $R^2$

- Continue until adjusted $R^2$ does not increase

---

## Model selection and interaction effects

If an interaction is included in the model, the main effects of both of
those variables must also be in the model

If a main effect is not in the model, then its interaction should not be 
in the model.

---

## Other model selection criteria

- Adjusted $R^2$ is one model selection criterion

- There are others out there (many many others!), we'll discuss some later in the 
course, and some you might see in another courses

---

class: center, middle

# Your turn

---

## Your turn

Work in teams on RStudio Cloud 

**Project:** Model selection for Paris Paintings

Make a copy and get started

---

## Planning

Decide on a subset of variables to consider for your analysis. 

- Think about it as focusing on certain aspects of the price determination, 
as opposed to all aspects. 

- You're allowed a maximum of 10 total variables to consider, including 
interactions. 

- The more variables you consider the longer model selection will take so 
keep that in mind.

---

## Planning (cont.)

Decide among these which variables might make sense to interact. Remember, 
we consider interactions when variables might behave differently for various 
levels of another variable. Ideally, you would get expert guidance for
choosing interactions. Below is a list of interactions compiled by them 
that might be potentially interesting:

.small[
- School of painting & landscape variables: `school_pntg` & `landsALL` / `lands_figs` / `lands_ment`
- Landscapes & paired paintings: `landsALL` / `lands_figs` / `lands_ment` & `paired`
- Other artists & paired paintings: `othartist` & `paired`
- Size & paired paintings: `surface` & `paired`
- Size & figures: `surface` & `figures` / `nfigures`
- Dealer & previous owner: `dealer` & `prevcoll`
- Winning bidder & prevcoll: `endbuyer` & `prevcoll`
- Winning bidder & artist living: `winningbiddertype` & `artistliving`
]

This is not an exhaustive list, so you might come up with others.

---

## Model fitting, selection, and interpretation

- Use backwards elimination to do model selection. Make sure to show 
each step of decision (though you don't have to interpret the models at 
each stage).

- Provide interpretations for the slopes for the final model you arrive at 
and create at least one visualization that supports your narrative.
