---
title: "Feature engineering"
subtitle: "<br><br> Data Science in a Box"
author: "[datasciencebox.org](https://datasciencebox.org/)"
output:
  xaringan::moon_reader:
    css: ["../xaringan-themer.css", "../slides.css"]
    lib_dir: libs
    anchor_sections: FALSE
    nature:
      ratio: "16:9"
      highlightLines: true
      highlightStyle: solarized-light
      countIncrementalSlides: false
---

```{r child = "../setup.Rmd"}
```

```{r packages, echo = FALSE, message=FALSE, warning=FALSE}
library(tidyverse)
library(tidymodels)
library(openintro)
library(lubridate)
library(knitr)
set.seed(1234)
options(
  warnPartialMatchArgs = FALSE,
  warnPartialMatchAttr = FALSE, 
  warnPartialMatchDollar = FALSE,
  width = 100
)
```

class: middle

# Feature engineering

---

## Feature engineering

- We prefer simple models when possible, but **parsimony** does not mean sacrificing accuracy (or predictive performance) in the interest of simplicity

--
- Variables that go into the model and how they are represented are just as critical to success of the model

--
- **Feature engineering** allows us to get creative with our predictors in an effort to make them more useful for our model (to increase its predictive performance) 

---

## Same training and testing sets as before

```{r}
# Fix random numbers by setting the seed 
# Enables analysis to be reproducible when random numbers are used 
set.seed(1116)

# Put 80% of the data into the training set 
email_split <- initial_split(email, prop = 0.80)

# Create data frames for the two sets:
train_data <- training(email_split)
test_data  <- testing(email_split)
```

---

## A simple approach: `mutate()`

```{r}
train_data %>%
  mutate(
    date = lubridate::date(time),
    dow  = wday(time),
    month = month(time)
    ) %>%
  select(time, date, dow, month) %>%
  sample_n(size = 5) # shuffle to show a variety
```

---

## Modeling workflow, revisited

- Create a **recipe** for feature engineering steps to be applied to the training data

--
- Fit the model to the training data after these steps have been applied

--
- Using the model estimates from the training data, predict outcomes for the test data

--
- Evaluate the performance of the model on the test data

---

class: middle

# Building recipes

---

## Initiate a recipe

```{r initiate-recipe, results="hide"}
email_rec <- recipe(
  spam ~ .,          # formula
  data = train_data  # data to use for cataloguing names and types of variables
  )

summary(email_rec)
```

.xsmall[
```{r echo=FALSE}
summary(email_rec) %>% print(n = 21)
```
]

---

## Remove certain variables

```{r}
email_rec <- email_rec %>%
  step_rm(from, sent_email)
```

.small[
```{r echo=FALSE}
email_rec
```
]

---

## Feature engineer date

```{r}
email_rec <- email_rec %>%
  step_date(time, features = c("dow", "month")) %>%
  step_rm(time)
```

.small[
```{r echo=FALSE}
email_rec
```
]

---

## Discretize numeric variables

```{r}
email_rec <- email_rec %>%
  step_cut(cc, attach, dollar, breaks = c(0, 1)) %>%
  step_cut(inherit, password, breaks = c(0, 1, 5, 10, 20))
```

.small[
```{r echo=FALSE}
email_rec
```
]

---

## Create dummy variables

```{r}
email_rec <- email_rec %>%
  step_dummy(all_nominal(), -all_outcomes())
```

.small[
```{r echo=FALSE}
email_rec
```
]

---

## Remove zero variance variables

Variables that contain only a single value

```{r}
email_rec <- email_rec %>%
  step_zv(all_predictors())
```

.small[
```{r echo=FALSE}
email_rec
```
]

---

## All in one place

```{r}
email_rec <- recipe(spam ~ ., data = email) %>%
  step_rm(from, sent_email) %>%
  step_date(time, features = c("dow", "month")) %>%               
  step_rm(time) %>%
  step_cut(cc, attach, dollar, breaks = c(0, 1)) %>%
  step_cut(inherit, password, breaks = c(0, 1, 5, 10, 20)) %>%
  step_dummy(all_nominal(), -all_outcomes()) %>%
  step_zv(all_predictors())
```

---

class: middle

# Building workflows

---

## Define model

```{r}
email_mod <- logistic_reg() %>% 
  set_engine("glm")

email_mod
```

---

## Define workflow

**Workflows** bring together models and recipes so that they can be easily applied to both the training and test data.

```{r}
email_wflow <- workflow() %>% 
  add_model(email_mod) %>% 
  add_recipe(email_rec)
```

.small[
```{r echo=FALSE}
email_wflow
```
]

---

## Fit model to training data

```{r}
email_fit <- email_wflow %>% 
  fit(data = train_data)
```

---

.small[
```{r}
tidy(email_fit) %>% print(n = 31)
```
]

---

## Make predictions for test data

```{r}
email_pred <- predict(email_fit, test_data, type = "prob") %>% 
  bind_cols(test_data) 

email_pred
```

---

## Evaluate the performance

.pull-left[
```{r roc, fig.show="hide"}
email_pred %>%
  roc_curve(
    truth = spam,
    .pred_1,
    event_level = "second"
  ) %>%
  autoplot()
```
]
.pull-right[
```{r ref.label="roc", echo=FALSE, out.width="100%"}
```
]

---

## Evaluate the performance

.pull-left[
```{r}
email_pred %>%
  roc_auc(
    truth = spam,
    .pred_1,
    event_level = "second"
  )
```
]
.pull-right[
```{r ref.label="roc", echo=FALSE, out.width="100%"}
```
]

---

class: middle

# Making decisions

---

## Cutoff probability: 0.5

.panelset[
.panel[.panel-name[Output]

Suppose we decide to label an email as spam if the model predicts the probability of spam to be **more than 0.5**.

```{r ref.label = "confusion-50,", echo = FALSE}
```
]
.panel[.panel-name[Code]
```{r confusion-50, results = "hide"}
cutoff_prob <- 0.5
email_pred %>%
  mutate(
    spam      = if_else(spam == 1, "Email is spam", "Email is not spam"),
    spam_pred = if_else(.pred_1 > cutoff_prob, "Email labelled spam", "Email labelled not spam")
    ) %>%
  count(spam_pred, spam) %>%
  pivot_wider(names_from = spam, values_from = n) %>%
  kable(col.names = c("", "Email is not spam", "Email is spam"))
```
]
]

---

## Cutoff probability: 0.25

.panelset[
.panel[.panel-name[Output]

Suppose we decide to label an email as spam if the model predicts the probability of spam to be **more than 0.25**.

```{r ref.label = "confusion-25,", echo = FALSE}
```
]
.panel[.panel-name[Code]
```{r confusion-25, results = "hide"}
cutoff_prob <- 0.25
email_pred %>%
  mutate(
    spam      = if_else(spam == 1, "Email is spam", "Email is not spam"),
    spam_pred = if_else(.pred_1 > cutoff_prob, "Email labelled spam", "Email labelled not spam")
    ) %>%
  count(spam_pred, spam) %>%
  pivot_wider(names_from = spam, values_from = n) %>%
  kable(col.names = c("", "Email is not spam", "Email is spam"))
```
]
]

---

## Cutoff probability: 0.75

.panelset[
.panel[.panel-name[Output]

Suppose we decide to label an email as spam if the model predicts the probability of spam to be **more than 0.75**.

```{r ref.label = "confusion-75,", echo = FALSE}
```
]
.panel[.panel-name[Code]
```{r confusion-75, results = "hide"}
cutoff_prob <- 0.75
email_pred %>%
  mutate(
    spam      = if_else(spam == 1, "Email is spam", "Email is not spam"),
    spam_pred = if_else(.pred_1 > cutoff_prob, "Email labelled spam", "Email labelled not spam")
    ) %>%
  count(spam_pred, spam) %>%
  pivot_wider(names_from = spam, values_from = n) %>%
  kable(col.names = c("", "Email is not spam", "Email is spam"))
```
]
]
