---
title: "Exam 01"
subtitle: "KEY"
output: 
  tufte::tufte_html:
    css: ../exam.css
    tufte_variant: "envisioned"
    highlight: pygments
    fig_height: 4
    fig_width: 7
link-citations: yes
---

## Academic Honesty Statement

**THIS IS AN INDIVIDUAL ASSESSMENT, THIS DOCUMENT AND YOUR ANSWERS ARE FOR YOUR 
EYES ONLY. ANY VIOLATION OF THIS POLICY WILL BE IMMEDIATELY REPORTED TO THE 
UNDERGRADUATE CONDUCT BOARD.**

*Replace the underscores below with your name acknowledging that you have read 
and understood the your institution's academic community standard.*

I, ____________, hereby state that I have not communicated with or gained 
information in any way from my classmates or anyone other than the Professor or 
TA during this exam, and that all work is my own.

## Packages

In this exam you will use, at a minimum, the tidyverse and nycflights13 packages. 
Load these packages on top of your document.

```{r load-packages, message=FALSE}
library(tidyverse)
library(nycflights13)
```

## Questions 

<!-- Question 1 -->

```{marginfigure}
5 points:  
- 2 pts - `count` OR `group_by` + `summarise`  
- either 1 pt - `arrange` + `desc` or `sort = TRUE` in `count` 
- 1 pt - `slice(1:10)`
- 1 pt narrative
```

1. What are the ten most common destinations for flights from NYC airports in 
2013? Your answer should be a table (or a data frame) with two columns 
(destination code and number of flights), arranged in descending order of number 
of flights. Also look up the name of the most common destination and state it 
in your narrative.

The most common destinations for flights from NYC is Chicago. The ten most 
common destinations are shown below.

```{r}
flights %>%
  count(dest) %>%
  arrange(desc(n)) %>%
  slice(1:10)
```

<!-- Question 2 -->

```{marginfigure}
5 points:  
- 1 pts - `count` OR `group_by` + `summarise`  
- either 1 pt - `arrange` + `desc` or `sort = TRUE` in `count` 
- 2 pt - `inner_join` with airlines  
- 1 pt narrative  
```

2. What are the top three airlines with the highest numbers of flights 
departing from NYC airports in 2013? Your answer should include a table 
(or a data frame) with three columns: carrier code, number of flights, 
and name of airline, and the airlines should be listed in descending order of 
number of flights. *Hint:* The `airlines` dataset can be used to look up 
airline names from their carrier codes.

The carrier with the highest number of flights departing from NYC airports in 
2013 is United Airlines, followed by JetBlue Airways and ExpressJet Airlines.

```{r}
flights %>%
  count(carrier) %>%
  arrange(desc(n)) %>%
  inner_join(airlines, by = "carrier")
```

<!-- Question 3 -->

```{marginfigure}
10 pts:  
- 2 pt - `filter` out `NAs` (or in `median`, `na.rm = TRUE`)  
- 2 pts - output sorted in order of median aarival delay  
- 2 pts - `inner_join` with `airlines`  
- 2 pts - Frontier Airlines, 6 minutes  
- 2 pts - Alaska Airlines, -17 minutes  
```

3. Consider only flights that have non-missing arrival delay information. 
Which carrier has the highest median arrival delay and which carrier has the 
lowest median arrival delay? Your answer should include a table (or a data 
frame) with carrier code, median arrival delay, and name of airline. 
**Hint:** First calculate the median arrival delay for all carriers, then 
join with `airlines` to match airline names. For comupational efficienc, avoid 
joining the large `flights` data frame directly with `airlines`.

Highest median arrival delay is on Frontier Airlines with 6 minutes, and 
lowest median arrival delay is on Alaska Airlines with -17 minutes (which means 
arriving 17 minutes earlier than scheduled).

```{r}
flights %>%
  filter(!is.na(arr_delay)) %>%
  group_by(carrier) %>%
  summarise(median_arr_delay = median(arr_delay)) %>%
  arrange(desc(median_arr_delay)) %>%
  inner_join(airlines, by = "carrier") %>%
  slice(c(1, n())) # optional
```

<!-- Question 4 -->

```{marginfigure}
10 pts:  
- 4 pts - JFK  
- 3 pts - Jan 9  
- 3 pts - 40.835 F  
```

4. Find the flight with the highest departure delay and determine on what day 
this happened. What was the mean temperature on that day at the origin 
airport of this flight? In your answer you should include carrier code and name, 
origin airport, departure date, as well as the mean temperature on that day 
at that airport. 

First, we find the flight with the highest departure dela and take a note 
of relevant information. The highest departure delay of 1301 minutes 
(approx 21.7 hours) was on a flight departing from JFK on Jan 9. 

```{r}
flights %>%
  arrange(desc(dep_delay)) %>%
  slice(1) %>%
  select(dep_delay, month, day, origin)
```

Then, we calculate the mean temperature on that day at the origin airport. 
The average temperature on this day was 40.8 F. `R`

```{r}
weather %>%
  filter(month == 1, day == 9, origin == "JFK") %>%
  summarise(mean_temp = mean(temp))
```

