---
title: "Coding style"
output:
  xaringan::moon_reader:
    css: "../slides.css"
    logo: img/sta199-logo-hex.png
    lib_dir: libs
    nature:
      highlightStyle: github
      highlightLines: true
      countIncrementalSlides: false
---

```{r packages, echo=FALSE, message=FALSE, warning=FALSE}
library(tidyverse)
```

```{r setup, include=FALSE}
# R options
options(
  htmltools.dir.version = FALSE, # for blogdown
  show.signif.stars = FALSE,     # for regression output
  warn = 1
  )
# Set dpi and height for images
knitr::opts_chunk$set(fig.height = 2.5, fig.width = 5, dpi = 300) 
# ggplot2 color palette with gray
color_palette <- list(gray = "#999999", 
                      salmon = "#E69F00", 
                      lightblue = "#56B4E9", 
                      green = "#009E73", 
                      yellow = "#F0E442", 
                      darkblue = "#0072B2", 
                      red = "#D55E00", 
                      purple = "#CC79A7")
htmltools::tagList(rmarkdown::html_dependency_font_awesome())
# For magick
dev.off <- function(){
  invisible(grDevices::dev.off())
}
# For ggplot2
ggplot2::theme_set(ggplot2::theme_bw())
```

class: center, middle

# Coding style

---

## Style guide

>"Good coding style is like correct punctuation: you can manage without it, butitsuremakesthingseasiertoread."
>
>Hadley Wickham

- Style guide for this course is based on the Tidyverse style guide: http://style.tidyverse.org/

- There's more to it than what we'll cover today, but we'll mention more as we introduce more functionality, and do a recap later in the semester

---

## File names and code chunk labels

- Do not use spaces in file names, use `-` or `_` to separate words
- Use all lowercase letters

```{r eval = FALSE}
# Good
ucb-admit.csv

# Bad
UCB Admit.csv
```

---

## Object names

- Use `_` to separate words in object names
- Use informative but short object names
- Do not reuse object names within an analysis

```{r eval = FALSE}
# Good
acs_employed

# Bad
acs.employed
acs2
acs_subset
acs_subsetted_for_males
```

---

## Spacing

- Put a space before and after all infix operators (=, +, -, <-, etc.), and when naming arguments in function calls. 
- Always put a space after a comma, and never before (just like in regular English).

```{r eval = FALSE}
# Good
average <- mean(feet / 12 + inches, na.rm = TRUE)

# Bad
average<-mean(feet/12+inches,na.rm=TRUE)
```

---

## ggplot

- Always end a line with `+`
- Always indent the next line

```{r eval = FALSE}
# Good
ggplot(diamonds, mapping = aes(x = price)) +
  geom_histogram()

# Bad
ggplot(diamonds,mapping=aes(x=price))+geom_histogram()
```

---

## Long lines

- Limit your code to 80 characters per line. This fits comfortably on a printed page with a reasonably sized font.

- Take advantage of RStudio editor's auto formatting for indentation at line breaks.

---

## Assignment

- Use `<-` not `=`

```{r eval = FALSE}
# Good
x <- 2

# Bad
x = 2
```

---

## Quotes

Use `"`, not `'`, for quoting text. The only exception is when the text already contains double quotes and no single quotes.

```{r eval = FALSE}
ggplot(diamonds, mapping = aes(x = price)) +
  geom_histogram() +
  # Good
  labs(title = "`Shine bright like a diamond`",
  # Good
       x = "Diamond prices",
  # Bad
       y = 'Frequency')
```
