---
title: "Building plots for various data types <br> `r emo::ji('bar_chart')``r emo::ji('bar_chart')`"
output:
  xaringan::moon_reader:
    css: "../slides.css"
    lib_dir: libs
    nature:
      ratio: "16:9"
      highlightLines: true
      highlightStyle: solarized-light
      countIncrementalSlides: false
---

```{r child = "../setup.Rmd"}
```

```{r packages, echo=FALSE, message=FALSE, warning=FALSE}
library(tidyverse)
library(datasauRus)
library(DT)
library(Tmisc)
library(dsbox)
```

class: center, middle

# Identifying variables

---

## Types of variables

- **Numerical variables** can be classified as **continuous** or **discrete** based on whether or not the variable can take on an infinite number of values or only non-negative whole numbers, respectively. 
- If the variable is **categorical**, we can determine if it is **ordinal** based on whether or not the levels have a natural ordering.

---

class: center, middle

# Visualizing numerical data

---

## Describing shapes of numerical distributions

* shape:
    * skewness: right-skewed, left-skewed, symmetric (skew is to the side of the longer tail)
    * modality: unimodal, bimodal, multimodal, uniform
* center: mean (`mean`), median (`median`), mode (not always useful)
* spread: range (`range`), standard deviation (`sd`), inter-quartile range (`IQR`)
* unusal observations

---

## Histograms

```{r out.width="75%", warning=FALSE}
ggplot(data = starwars, mapping = aes(x = height)) +
  geom_histogram(binwidth = 10)
```

---

## Density plots

```{r out.width="75%", warning = FALSE}
ggplot(data = starwars, mapping = aes(x = height)) +
  geom_density()
```

---

## Box plots

```{r out.width="60%", warning = FALSE}
ggplot(data = starwars, mapping = aes(y = height)) +
  geom_boxplot()
```

---

class: center, middle

# Visualizing relationships between numerical and categorical data

---

## Side-by-side box plots

```{r out.width="75%", warning=FALSE}
ggplot(data = starwars, mapping = aes(y = height, x = gender)) +
  geom_boxplot()
```

---

## Scatter plot...

This is not a great representation of these data.

```{r out.width="60%", warning=FALSE}
ggplot(data = starwars, mapping = aes(y = height, x = gender)) +
  geom_point()
```

---

## Violin plots

```{r out.width="75%", warning=FALSE}
ggplot(data = starwars, mapping = aes(y = height, x = gender)) +
  geom_violin()
```

---

## Jitter plot

```{r echo=FALSE}
set.seed(1234)
```


```{r out.width="75%", warning=FALSE}
ggplot(data = starwars, mapping = aes(y = height, x = gender)) +
  geom_jitter()
```

---

## Beeswarm plots

```{r out.width="75%", warning = FALSE}
library(ggbeeswarm)
ggplot(data = starwars, mapping = aes(y = height, x = gender)) +
  geom_beeswarm()
```

---

class: center, middle

# Visualizing categorical data

---

## Bar plots

```{r out.width="80%"}
ggplot(data = starwars, mapping = aes(x = gender)) +
  geom_bar()
```

---

## Segmented bar plots, counts

```{r out.width="80%"}
ggplot(data = starwars, mapping = aes(x = gender, fill = hair_color)) +
  geom_bar()
```

---

## Recode hair color

Using the `fct_other()` function from the **forcats** package, which is also part of the **tidyverse**.

```{r}
starwars <- starwars %>%
  mutate(hair_color2 = 
           fct_other(hair_color, 
                     keep = c("black", "brown", "brown", "blond")
                     )
         )
```

---

## Segmented bar plots, counts

```{r out.width="70%"}
ggplot(data = starwars, 
       mapping = aes(x = gender, fill = hair_color2)) +
  geom_bar() +
  coord_flip()
```

---

## Segmented bar plots, proportions

```{r out.width="70%"}
ggplot(data = starwars, 
       mapping = aes(x = gender, fill = hair_color2)) +
  geom_bar(position = "fill") +
  coord_flip()
  labs(y = "proportion")
```

---

.question[
Which bar plot is a more useful representation for visualizing the relationship between gender and hair color?
]

```{r echo=FALSE, out.width="50%"}
ggplot(data = starwars, mapping = aes(x = gender, fill = hair_color2)) +
  geom_bar() +
  coord_flip()
```

```{r echo=FALSE, out.width="50%"}
ggplot(data = starwars, mapping = aes(x = gender, fill = hair_color2)) +
  geom_bar(position = "fill") +
  coord_flip() +
  labs(y = "proportion")
```

---

class: center, middle

# Why do we visualize?

---

## Data: `datasaurus_dozen`

Below is an exceprt from the `datasaurus_dozen` dataset:

```{r datasaurus-view-wide, echo=FALSE}
library(datasauRus)
datasaurus_dozen_wide %>%
  select(1:8)
```

---

## Summary statistics

.small[
```{r datasaurus-cor}
datasaurus_dozen %>%
  group_by(dataset) %>%
  summarise(r = cor(x, y))
```
]

---

## 

.question[
How similar do the relationships between `x` and `y` in the thirteen datasets look?
How similar are they based on summary stats?
]

```{r datasaurus-plot, echo=FALSE, fig.height=4, fig.width=7}
ggplot(datasaurus_dozen, aes(x = x, y = y, color = dataset))+
  geom_point()+
  facet_wrap(~ dataset, ncol = 5) +
  theme(legend.position = "none")
```

---

## Anscombe's quartet

```{r quartet-for-show, eval=FALSE}
library(Tmisc)
quartet
```

.pull-left[
```{r quartet-view1, echo=FALSE}
quartet[1:22,]
```
]
.pull-right[
```{r quartet-view2, echo=FALSE}
quartet[23:44,]
```
]

---

## Summarising Anscombe's quartet

```{r quartet-summary}
quartet %>%
  group_by(set) %>%
  summarise(
    mean_x = mean(x), 
    mean_y = mean(y),
    sd_x = sd(x),
    sd_y = sd(y),
    r = cor(x, y)
  )
```

---

## Visualizing Anscombe's quartet

```{r quartet-plot, out.width="75%"}
ggplot(quartet, aes(x = x, y = y)) +
  geom_point() +
  facet_wrap(~ set, ncol = 4)
```

---

## Age at first kiss

.question[
Do you see anything out of the ordinary?
]

```{r fig.height=2.5, echo=FALSE, warning=FALSE}
ggplot(student_survey, aes(x = first_kiss)) +
  geom_histogram(binwidth = 1) +
  labs(title = "How old were you when you had your first kiss?")
```

---

## Facebook visits

.question[
How are people reporting lower vs. higher values of FB visits?
]

```{r fig.height=2.5, echo=FALSE, warning=FALSE}
ggplot(student_survey, aes(x = fb_visits_per_day)) +
  geom_histogram(binwidth = 1) +
  labs(title = "How many times do you go on Facebook per day?", x = "")
```
