---
title: "Building plots for various data types <br> `r emo::ji('bar_chart')`"
author: ""
output:
  xaringan::moon_reader:
    css: "../slides.css"
    lib_dir: libs
    nature:
      ratio: "16:9"
      highlightLines: true
      highlightStyle: solarized-light
      countIncrementalSlides: false
---

```{r child = "../setup.Rmd"}
```

```{r packages, echo=FALSE, message=FALSE, warning=FALSE}
library(emo)
library(tidyverse)
library(fontawesome)
library(magick)
```

class: middle

# Identifying variables

---

## Number of variables involved

- Univariate data analysis - distribution of single variable
- Bivariate data analysis - relationship between two variables
- Multivariate data analysis - relationship between many variables at once, usually focusing on the relationship between two while conditioning for others

---

## Types of variables

- **Numerical variables** can be classified as **continuous** or **discrete** based on whether or not the variable can take on an infinite number of values or only non-negative whole numbers, respectively.
- If the variable is **categorical**, we can determine if it is **ordinal** based on whether or not the levels have a natural ordering.

---

class: middle

# Visualizing numerical data

---

## Describing shapes of numerical distributions

- shape:
    - skewness: right-skewed, left-skewed, symmetric (skew is to the side of the longer tail)
    - modality: unimodal, bimodal, multimodal, uniform
- center: mean (`mean`), median (`median`), mode (not always useful)
- spread: range (`range`), standard deviation (`sd`), inter-quartile range (`IQR`)
- unusal observations

---

## Histograms

```{r out.width="75%"}
ggplot(data = starwars, mapping = aes(x = height)) +
  geom_histogram(binwidth = 10)
```

---

## Density plots

```{r out.width="75%"}
ggplot(data = starwars, mapping = aes(x = height)) +
  geom_density()
```

---

class: middle

# Visualizing categorical data

---

## Bar plots

```{r out.width="80%"}
ggplot(data = starwars, mapping = aes(x = gender)) +
  geom_bar()
```

---

## Segmented bar plots, counts

```{r out.width="80%"}
ggplot(data = starwars, mapping = aes(x = gender, fill = hair_color)) +
  geom_bar()
```

---

## Recode hair color

```{r}
starwars <- starwars %>%
  mutate(hair_color2 = 
           fct_other(hair_color, 
                     keep = c("black", "brown", "brown", "blond")
                     )
         )
```

---

## Segmented bar plots, counts

```{r out.width="70%"}
ggplot(data = starwars, 
       mapping = aes(x = gender, fill = hair_color2)) +
  geom_bar() +
  coord_flip()
```

---

## Segmented bar plots, proportions

```{r out.width="70%"}
ggplot(data = starwars, 
       mapping = aes(x = gender, fill = hair_color2)) +
  geom_bar(position = "fill") +
  coord_flip()
  labs(y = "proportion")
```

---

.question[
Which bar plot is a more useful representation for visualizing the relationship between gender and hair color?
]

```{r echo=FALSE, out.width="50%"}
ggplot(data = starwars, mapping = aes(x = gender, fill = hair_color2)) +
  geom_bar() +
  coord_flip()
```

```{r echo=FALSE, out.width="50%"}
ggplot(data = starwars, mapping = aes(x = gender, fill = hair_color2)) +
  geom_bar(position = "fill") +
  coord_flip() +
  labs(y = "proportion")
```

---

.your-turn[

Time to actually play around with the Star Wars dataset!

- Go to RStudio Cloud ([rstd.io/dsbox-cloud](http://rstd.io/dsbox-cloud)) and start `AE 03 - StarWars + Data visualization`.
- Open the R Markdown document and complete the exercise (and if time allows, the stretch goal exercise).
- Once done, place a green sticky on your laptop. If you have questions, place a pink sticky. ]

---

class: middle

# Visualizing relationships between numerical and categorical data

---

## Side-by-side box plots

```{r out.width="75%", warning=FALSE}
ggplot(data = starwars, mapping = aes(y = height, x = gender)) +
  geom_boxplot()
```

---

## Scatter plot...

This is not a great representation of these data.

```{r out.width="60%", warning=FALSE}
ggplot(data = starwars, mapping = aes(y = height, x = gender)) +
  geom_point()
```

---

## Violin plots

```{r out.width="75%", warning=FALSE}
ggplot(data = starwars, mapping = aes(y = height, x = gender)) +
  geom_violin()
```

---

## Jitter plot

```{r echo=FALSE}
set.seed(1234)
```

```{r out.width="75%", warning=FALSE}
ggplot(data = starwars, mapping = aes(y = height, x = gender)) +
  geom_jitter()
```

---

## Beeswarm plots

```{r out.width="75%", warning = FALSE}
library(ggbeeswarm)
ggplot(data = starwars, mapping = aes(y = height, x = gender)) +
  geom_beeswarm()
```

---

## Ridge plots

```{r out.width="70%", warning = FALSE}
library(ggridges)
ggplot(starwars, aes(x = height, y = gender, fill = gender, color = gender)) + 
  geom_density_ridges(alpha = 0.5)
```

---

class: middle

# Visualizing relationships between numerical and numerical data

---

## Scatterplot

```{r}
starwars_without_jabba <- starwars %>% filter(mass < 500)
```

```{r out.width="75%", warning=FALSE}
ggplot(starwars_without_jabba, aes(x = height, y = mass)) +
  geom_point()
```

---

## Hex plots

Especially useful for large data (which this one is not!)

```{r out.width="75%", warning=FALSE}
ggplot(starwars_without_jabba, aes(x = height, y = mass)) +
  geom_hex()
```
