---
title: "UN Votes"
author: "Mine Çetinkaya-Rundel"
date: "`r Sys.Date()`"
output: 
  html_document: 
    toc: yes
    toc_float: yes
---

## Introduction

How do various countries vote in the United Nations General Assembly, how have 
their voting patterns evolved throughout time, and how similarly or differently 
do they view certain issues? Answering these questions (at a high level) is the 
focus of this analysis.

### Packages

We will use the __tidyverse__, __lubridate__, and __scales__ packages for the 
data wrangling and visualization, and the __DT__ package for interactive display 
of tabular output.

```{r load-packages, warning=FALSE, message=FALSE}
library(tidyverse)
library(lubridate)
library(scales)
library(DT)
```

### Data

The data we're using originally come from the **unvotes** package, but it's been 
modified a bit (by joining the various data frames provided in the package) to 
help you get started with the analysis.

```{r load-data}
unvotes <- read_rds("data/unvotes.rds")
```


## UN voting patterns {#voting}

Let's  create a data visualization that displays how the voting record of the 
UK & NI changed over time on a variety of issues, and compares it 
to two other countries: US and Turkey.

We can easily change which countries are being plotted by changing which 
countries the code above `filter`s for. Note that the country name should be 
spelled and capitalized exactly the same way as it appears in the data. See 
the [Appendix](#appendix) for a list of the countries in the data.

```{r plot-yearly-yes-issue, fig.width=10, fig.height=6, message=FALSE}
unvotes %>%
  filter(country %in% c("UK & NI", "US", "Turkey")) %>%
  mutate(year = year(date)) %>%
  group_by(country, year, issue) %>%
  summarize(percent_yes = mean(vote == "yes")) %>%
  ggplot(mapping = aes(x = year, y = percent_yes, color = country)) +
  geom_point(alpha = 0.4) +
  geom_smooth(method = "loess", se = FALSE) +
  facet_wrap(~issue) +
  scale_y_continuous(labels = percent) +
  labs(
    title = "Percentage of 'Yes' votes in the UN General Assembly",
    subtitle = "1946 to 2015",
    y = "% Yes",
    x = "Year",
    color = "Country"
  )
```


## References {#references}

1. David Robinson (2017). [unvotes](https://CRAN.R-project.org/package=unvotes): 
   United Nations General Assembly Voting Data. R package version 0.2.0.
1. Erik Voeten "Data and Analyses of Voting in the UN General Assembly" 
   Routledge Handbook of International Organization, edited by Bob Reinalda 
   (published May 27, 2013).
1. Much of the analysis has been modeled on the examples presented in the 
   [unvotes package vignette](https://cran.r-project.org/web/packages/unvotes/vignettes/unvotes.html).


## Appendix {#appendix}

Below is a list of countries in the dataset:

```{r list-countries, echo=FALSE}
unvotes %>% 
  select(country) %>%
  arrange(country) %>% 
  distinct() %>%
  datatable()
```
