---
title: "The Office"
author: "Mine Çetinkaya-Rundel"
output: github_document
---

```{r load-packages, message = FALSE}
library(tidyverse)
library(tidymodels)
library(schrute)
library(lubridate)
```

Use `theoffice` data from the [**schrute**](https://bradlindblad.github.io/schrute/) package to predict IMDB scores for episodes of The Office.

```{r}
glimpse(theoffice)
```

Fix `air_date` for later use.

```{r}
theoffice <- theoffice %>%
  mutate(air_date = ymd(as.character(air_date)))
```

We will

-   engineer features based on episode scripts
-   train a model
-   perform cross validation
-   make predictions

Note: The episodes listed in `theoffice` don't match the ones listed in the data we used in the [cross validation lesson](https://ids-s1-20.github.io/slides/week-10/w10-d02-cross-validation/w10-d02-cross-validation.html).

```{r}
theoffice %>%
  distinct(season, episode)
```

### Exercise 1 - Calculate the percentage of lines spoken by Jim, Pam, Michael, and Dwight for each episode of The Office.

```{r lines}

```

### Exercise 2 - Identify episodes that touch on Halloween, Valentine's Day, and Christmas.

```{r special-episodes}

```

### Exercise 3 - Put together a modeling dataset that includes features you've engineered. Also add an indicator variable called `michael` which takes the value `1` if Michael Scott (Steve Carrell) was there, and `0` if not. Note: Michael Scott (Steve Carrell) left the show at the end of Season 7.

```{r office-df}

```

### Exercise 4 - Split the data into training (75%) and testing (25%).

```{r split}
set.seed(1122)

```

### Exercise 5 - Specify a linear regression model.

```{r model}

```

### Exercise 6 - Create a recipe that updates the role of `episode_name` to not be a predictor, removes `air_date` as a predictor, uses `season` as a factor, and removes all zero variance predictors.

```{r recipe}

```

### Exercise 7 - Build a workflow for fitting the model specified earlier and using the recipe you developed to preprocess the data.

```{r workflow}

```

### Exercise 8 - Fit the model to training data and interpret a couple of the slope coefficients.

```{r fit}

```

### Exercise 9 - Perform 5-fold cross validation and view model performance metrics.

```{r cv, message=FALSE, error = TRUE}
set.seed(345)
folds <- vfold_cv(___, v = ___)
folds

set.seed(456)
office_fit_rs <- ___ %>%
  ___(___)

___(office_fit_rs)
```

### Exercise 10 - Use your model to make predictions for the testing data and calculate the RMSE. Also use the model developed in the [cross validation lesson](https://ids-s1-20.github.io/slides/week-10/w10-d02-cross-validation/w10-d02-cross-validation.html) to make predictions for the testing data and calculate the RMSE as well. Which model did a better job in predicting IMDB scores for the testing data?

#### New model

```{r new-model}

```

#### Old model

```{r old-model, error = TRUE}
office_mod_old <- linear_reg() %>%
  set_engine("lm")

office_rec_old <- recipe(imdb_rating ~ season + episode + total_votes + air_date, data = office_train) %>%
  # extract month of air_date
  step_date(air_date, features = "month") %>%
  step_rm(air_date) %>%
  # make dummy variables of month 
  step_dummy(contains("month")) %>%
  # remove zero variance predictors
  step_zv(all_predictors())

office_wflow_old <- workflow() %>%
  add_model(office_mod_old) %>%
  add_recipe(office_rec_old)

office_fit_old <- office_wflow_old %>%
  fit(data = office_train)

tidy(office_fit_old)

___
```
